---
title: "Analysis Script for Social and Cultural Dynamics 2026"
author: "Andrea Wiwe Kjær-Hansen, Mikkel Krøjer Svendsen"
format: html
---

# R setup
```{r, message=FALSE}
# Packages
library(rethinking)
library(patchwork)
library(tidyverse)
library(kableExtra)

# Theme settings
theme_set(theme_minimal())
color1 <- "#564787"
color2 <- "#F3C969"

# data directory
source("data_dir.R")
```

```{r}
# read files helper
read_files <- function(pattern) {
  list.files(
    data_dir,
    pattern = pattern,
    full.names = TRUE
  ) |>
    map_dfr(~ read.delim2(.x, sep = ","))
}
```

```{r}
# read files
d_simon <- read_files("simon") |>
  mutate(
    rt_ms = as.numeric(ifelse(rt_ms == "", NA, rt_ms)),
    response_key = ifelse(response_key == "", NA, response_key),
    side_congruent = (shape == "square" & side == "left") |
      (shape == "diamond" & side == "right")
  )

d_tapping <- read_files("tapping")

d_info <- read_files("group_info") |>
  mutate(
    gender = ifelse(dyad_id == 4 & participant == 2, "Male", gender) # one participant entered wrong gender
  )
```

# Descriptive statistics and data cleaning
## Participants
```{r}
age <- d_info |>
  summarize(
    age_mean = mean(age),
    age_sd = sd(age)
  )

gender <- d_info |>
  count(gender) |>
  mutate(
    prop = n / sum(n)
  )

condition <- d_tapping |>
  distinct(dyad_id, .keep_all = T) |>
  count(condition, name = "n_dyads")

print(age)
print(gender)
print(condition)
```

## Calculating the Social Simon Effect (SSE)
First, add condition column to data
```{r}
# add condition to d_simon
d_simon <- d_simon |>
  left_join(d_tapping |> distinct(dyad_id, condition), by = "dyad_id")

str(d_simon)
```

We only look at correct go trials (`responded == 1` and `correct == 1`). We filter out the non-response participant from each trial.
```{r}
d <- d_simon |>
  filter(responded == 1 & correct == 1) |>
  mutate(participant_id = paste(dyad_id, participant, sep = "_")) |>
  filter(
    (participant == 1 & shape == "square") |
      (participant == 2 & shape == "diamond")
  )

str(d)
```

RT summary by condition x task type x congruency

NOTE: If knitting to Word rather than PDF, swap `kable_styling(latex_options = ...)` for `kable_styling(bootstrap_options = "striped")` since the LaTeX options are PDF-only.
```{r}
rt_summary <- bind_rows(
  # Individual: collapse across condition
  d |>
    filter(task == "individual") |>
    group_by(task, side_congruent) |>
    summarise(
      mean_rt = mean(rt_ms, na.rm = TRUE),
      sd_rt = sd(rt_ms, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    ) |>
    mutate(condition = "Not yet assigned"),

  # Joint: keep condition
  d |>
    filter(task == "joint") |>
    group_by(task, condition, side_congruent) |>
    summarise(
      mean_rt = mean(rt_ms, na.rm = TRUE),
      sd_rt = sd(rt_ms, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    )
) |>
  select(task, condition, side_congruent, mean_rt, sd_rt, n)
```

```{r}
# Condition-split rows
condition_split <- d |>
  group_by(task, condition, side_congruent) |>
  summarise(
    mean_rt = mean(rt_ms, na.rm = TRUE),
    sd_rt = sd(rt_ms, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = side_congruent, values_from = c(mean_rt, sd_rt)) |>
  rename(
    congruent = mean_rt_TRUE,
    incongruent = mean_rt_FALSE,
    sd_con = sd_rt_TRUE,
    sd_incon = sd_rt_FALSE
  ) |>
  mutate(congruency_effect = incongruent - congruent)

# Pooled individual row
pooled_individual <- d |>
  filter(task == "individual") |>
  group_by(task, side_congruent) |>
  summarise(
    mean_rt = mean(rt_ms, na.rm = TRUE),
    sd_rt = sd(rt_ms, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = side_congruent, values_from = c(mean_rt, sd_rt)) |>
  rename(
    congruent = mean_rt_TRUE,
    incongruent = mean_rt_FALSE,
    sd_con = sd_rt_TRUE,
    sd_incon = sd_rt_FALSE
  ) |>
  mutate(congruency_effect = incongruent - congruent, condition = "(pooled)")

merged_summary <- bind_rows(pooled_individual, condition_split) |>
  select(
    task,
    condition,
    congruent,
    sd_con,
    incongruent,
    sd_incon,
    congruency_effect
  )

# Table
table1 <- merged_summary |>
  mutate(
    condition = replace_na(condition, "-"),
    congruent = paste0(
      formatC(congruent, format = "f", digits = 1),
      " (",
      formatC(sd_con, format = "f", digits = 1),
      ")"
    ),
    incongruent = paste0(
      formatC(incongruent, format = "f", digits = 1),
      " (",
      formatC(sd_incon, format = "f", digits = 1),
      ")"
    ),
    congruency_effect = formatC(congruency_effect, format = "f", digits = 1)
  ) |>
  select(task, condition, congruent, incongruent, congruency_effect) |>
  rename(
    Task = task,
    Condition = condition,
    `Congruent` = congruent,
    `Incongruent` = incongruent,
    `Congruency effect` = congruency_effect
  ) |>
  kbl(
    caption = "Table 1. Mean RT in ms (SD) and congruency effect (incongruent - congruent) by task and condition.",
    align = c("l", "l", "r", "r", "r"),
    booktabs = TRUE
  ) |>
  kable_styling(
    latex_options = c("striped", "hold_position"),
    full_width = FALSE
  ) |>
  collapse_rows(columns = 1, valign = "top") |>
  add_header_above(c(" " = 2, "Mean RT ms (SD)" = 2, " " = 1)) |>
  footnote(
    general = "Individual trials were completed prior to the synchrony manipulation. Condition labels for individual rows reflect the condition the dyad was later assigned to, not an active manipulation.",
    general_title = "Note:",
    footnote_as_chunk = TRUE
  )

table1
```

The asynchrony condition looks as expected. There's a small congruency effect in the individual task (0.6 ms) and a larger one in the joint task (18 ms). BUT in the synchrony condition, individuals are slower for congruent trials than incongruent trials. This may just be a noisy grand average... worth investigating.

SSE: difference in congruency effect between joint and individual.
```{r}
congruency_effect <- d |>
  group_by(condition, task, side_congruent) |>
  summarise(mean_rt = mean(rt_ms, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = side_congruent, values_from = mean_rt) |>
  rename(congruent = `TRUE`, incongruent = `FALSE`) |>
  mutate(congruency_effect = incongruent - congruent)

sse_summary <- congruency_effect |>
  select(condition, task, congruency_effect) |>
  pivot_wider(names_from = task, values_from = congruency_effect) |>
  mutate(SSE = joint - individual)

table2 <- sse_summary |>
  mutate(
    condition = replace_na(condition, "-"),
    across(c(individual, joint, SSE), ~ formatC(.x, format = "f", digits = 1))
  ) |>
  rename(
    Condition = condition,
    Individual = individual,
    Joint = joint,
    SSE = SSE
  ) |>
  kbl(
    caption = "Table 2. Congruency effects (ms) by task and Social Simon Effect (SSE = joint - individual).",
    align = c("l", "r", "r", "r"),
    booktabs = TRUE
  ) |>
  kable_styling(
    latex_options = c("striped", "hold_position"),
    full_width = FALSE
  ) |>
  add_header_above(c(" " = 1, "Congruency effect (ms)" = 2, " " = 1))

table2
```

SSE *is* slightly larger for sync than async, though not by a lot.  

# Visualizations
## RT distributions by condition x task type x congruency
```{r}
d |>
  mutate(
    task = factor(
      task,
      levels = c("individual", "joint"),
      labels = c("Individual", "Joint")
    ),
    congruency = factor(
      side_congruent,
      levels = c(FALSE, TRUE),
      labels = c("Incongruent", "Congruent")
    ),
    condition = factor(
      condition,
      levels = c("asynchrony", "synchrony"),
      labels = c("Asynchrony", "Synchrony")
    )
  ) |>
  filter(rt_ms < 2000) |>
  ggplot(aes(x = rt_ms, fill = congruency)) +
  geom_density(alpha = .5) +
  facet_grid(condition ~ task) +
  scale_fill_manual(values = c(color1, color2)) +
  labs(
    title = "RT distributions by condition, task type, and congruency",
    x = "RT (ms)",
    y = "Density",
    fill = ""
  )
```

SSE is slightly visible in the asynchrony/joint cell but not elsewhere... maybe in synchrony/joint but very noisy.  

Individual (go/no-go) task panels show minimal congruency separation in both conditions which is expected. The congruency manipulation is only meaningful in the joint context.  

So, data appear to show a reduced or reversed SSE in the synchrony condition relative to asynchrony contrary to expectation?  

## Individual dyad trajectories 
```{r}
sse_plot_data <- d |>
  group_by(condition, task, dyad_id, side_congruent) |>
  summarise(mean_rt = mean(rt_ms, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = side_congruent, values_from = mean_rt) |>
  rename(incongruent = `FALSE`, congruent_rt = `TRUE`) |>
  mutate(
    sse = incongruent - congruent_rt,
    task = factor(
      task,
      levels = c("individual", "joint"),
      labels = c("Individual", "Joint")
    ),
    condition = factor(
      condition,
      levels = c("asynchrony", "synchrony"),
      labels = c("Asynchrony", "Synchrony")
    )
  )
```

```{r}
sse_plot_data |>
  ggplot(aes(
    x = task,
    y = sse,
    colour = condition,
    group = interaction(dyad_id, condition)
  )) +
  geom_line(alpha = .4, position = position_dodge(.2)) +
  geom_point(alpha = .4, position = position_dodge(.2)) +
  stat_summary(
    aes(group = condition),
    fun = mean,
    geom = "line",
    linewidth = 1.2,
    position = position_dodge(.2)
  ) +
  stat_summary(
    aes(group = condition),
    fun = mean,
    geom = "point",
    size = 3,
    position = position_dodge(.2)
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_colour_manual(values = c(color1, color2)) +
  labs(
    title = "Figure 1. Congruency effect per dyad by task type and synchrony condition",
    subtitle = "Thin lines: individual dyads\nThick lines: group mean",
    x = "",
    y = "Congruency effect (ms)\nIncongruent - Congruent",
    colour = ""
  )
```

Individual task variance. Some show positive congruency effects, some negative. This is expected noise.  

The joint task is more consistent. Nearly all dyads show positive congruency effects in the joint task. We're replicating the SSE reliably, nice. 

# Data for analysis
`ulam()` requires lists rather than data frames.
```{r}
d <- d |>
  mutate(
    congruent = as.integer(side_congruent), # 0 = incongruent, 1 = congruent
    joint = as.integer(task == "joint"), # 0 = individual, 1 = joint
    sync = as.integer(condition == "synchrony"), # 0 = async, 1 = sync
    dyad_id = as.integer(as.factor(dyad_id)),
    participant_id = as.integer(as.factor(participant_id))
  )

dat <- list(
  rt = d$rt_ms,
  congruent = d$congruent,
  joint = d$joint,
  sync = d$sync,
  dyad = d$dyad_id,
  pid = d$participant_id
)
```

# Analysis
## Model 0 - intercept only
### Prior predictive
```{r}
n_samples <- 2000
prior_mu <- rnorm(n_samples, mean = 6.1, sd = .2)
prior_sigma <- rexp(n_samples, rate = 3)
prior_rt <- rlnorm(n_samples, meanlog = prior_mu, sdlog = prior_sigma)

# Plot
tibble(rt = prior_rt) |>
  filter(rt < 5000) |> # trim extreme tails for visualisation
  ggplot(aes(x = rt)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check - Model 0",
    subtitle = "Red dashed lines: plausible RT range (150 ms - 1800 ms)",
    x = "RT (ms)",
    y = "Density"
  )
```

### Model fit
```{r}
m0 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a,
    a ~ dnorm(6.1, .2),
    sigma ~ dexp(3)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  log_lik = TRUE,
  seed = 42
)
```

```{r}
precis(m0)
```

### Posterior predictive check
sim() draws from the posterior predictive distribution (includes sigma)
link() draws mu only. Use sim() here since we want full RT distributions:
```{r}
post0 <- extract.samples(m0)

rt_sim0 <- sim(m0, n = n_samples)

# One simulated RT per observation per posterior draw
# colMeans gives the average posterior predictive per observation,
# but for a density plot we want the full distribution
# so flatten the entire matrix instead

rt_ppc0 <- as.vector(rt_sim0)

tibble(
  rt     = c(dat$rt, rt_ppc0),
  source = c(rep("Observed", length(dat$rt)),
             rep("Posterior predictive", length(rt_ppc0)))
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = .5) +
  labs(
    title = "Posterior predictive check — Model 0",
    x = "RT (ms)", y = "Density", fill = ""
  )
```

### Prior-posterior comparison
```{r}
n_post <- length(post0$a)

prior_post0 <- tibble(
  prior__a = rnorm(n_post, 6.1, .2),
  posterior__a = post0$a,
  prior__sigma = rexp(n_post, 3),
  posterior__sigma = post0$sigma
) |>
  pivot_longer(everything(),
               names_to  = c("source", "parameter"),
               names_sep = "__")

# Compute posterior range per parameter for zooming
zoom <- prior_post0 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001),
            xmax = quantile(value, 0.999))

prior_post0 |>
  left_join(zoom, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = .5) +
  facet_wrap(~ parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison - Model 0",
    x = "", y = "Density", fill = ""
  )
```

## Model 1 - main effects
### Prior predictive
```{r}
n_post <- 2000

prior_a <- rnorm(n_post, 6.1, .2)
prior_b_cong <- rnorm(n_post, 0, .05)
prior_b_joint <- rnorm(n_post, -.05, .05)
prior_b_sync <- rnorm(n_post, 0, .05)
prior_sigma <- rexp(n_post, 3)

# Compute prior mu for each combination of predictors (reference cell: 0/0/0)
prior_mu <- prior_a +
  prior_b_cong * 1 + # congruent
  prior_b_joint * 1 + # joint
  prior_b_sync * 1 # sync

prior_rt <- rlnorm(n_post, meanlog = prior_mu, sdlog = prior_sigma)

tibble(rt = prior_rt) |>
  filter(rt < 5000) |>
  ggplot(aes(x = rt)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check - Model 1",
    subtitle = "Red dashed lines: plausible RT range (150-1800 ms)",
    x = "RT (ms)",
    y = "Density"
  )
```

### Model fit
```{r}
m1 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong * congruent +
      b_joint * joint +
      b_sync * sync,
    a ~ dnorm(6.1, .2),
    b_cong ~ dnorm(0, .05),
    b_joint ~ dnorm(-.05, .05),
    b_sync ~ dnorm(0, .05),
    sigma ~ dexp(3)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  log_lik = TRUE,
  seed = 42
)
```

```{r}
precis(m1)
```

### Posterior predictive check
```{r}
post1 <- extract.samples(m1)
rt_sim1 <- sim(m1)
rt_ppc1 <- as.vector(rt_sim1)

tibble(
  rt = c(dat$rt, rt_ppc1),
  source = c(
    rep("Observed", length(dat$rt)),
    rep("Posterior predictive", length(rt_ppc1))
  )
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = .5) +
  labs(
    title = "Posterior predictive check — Model 1",
    x = "RT (ms)",
    y = "Density",
    fill = ""
  )
```

### Prior-posterior comparison
```{r}
n_post <- length(post1$a)

prior_post1 <- tibble(
  prior__a = rnorm(n_post, 6.1, .2),
  posterior__a = post1$a,
  prior__cong = rnorm(n_post, 0, .05),
  posterior__cong = post1$b_cong,
  prior__joint = rnorm(n_post, -.05, .05),
  posterior__joint = post1$b_joint,
  prior__sync = rnorm(n_post, 0, .05),
  posterior__sync = post1$b_sync,
  prior__sigma = rexp(n_post, 3),
  posterior__sigma = post1$sigma
) |>
  pivot_longer(
    everything(),
    names_to = c("source", "parameter"),
    names_sep = "__"
  )

zoom1 <- prior_post1 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001), xmax = quantile(value, 0.999))

prior_post1 |>
  left_join(zoom1, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = .5) +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison - Model 1",
    x = "",
    y = "Density",
    fill = ""
  )
```

## Model 2 - main effects + varying intercepts
### Prior predictive
```{r}
n_post <- 2000

prior_a <- rnorm(n_post, 6.1, .2)
prior_b_cong <- rnorm(n_post, 0, .05)
prior_b_joint <- rnorm(n_post, -.05, 0.05)
prior_b_sync <- rnorm(n_post, 0, .05)
prior_sigma <- rexp(n_post, 3)
prior_sigma_pid <- rexp(n_post, 5)
prior_sigma_dyad <- rexp(n_post, 5)

# Simulate varying intercepts
prior_u_pid <- rnorm(n_post, 0, prior_sigma_pid)
prior_u_dyad <- rnorm(n_post, 0, prior_sigma_dyad)

prior_mu <- prior_a +
  prior_b_cong * 1 +
  prior_b_joint * 1 +
  prior_b_sync * 1 +
  prior_u_pid +
  prior_u_dyad

prior_rt <- rlnorm(n_post, meanlog = prior_mu, sdlog = prior_sigma)

tibble(rt = prior_rt) |>
  filter(rt < 5000) |>
  ggplot(aes(x = rt)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check - Model 2",
    subtitle = "Red dashed lines: plausible RT range (150-1800 ms)",
    x = "RT (ms)",
    y = "Density"
  )
```

This distribution is wider than m0/m1 if variance components kept at `rexp(3)`: that would make the right tail heavier and push more mass below 150 ms. This is expected since the two random effect terms (`u_pid` and `u_dyad`) each add variance on the log scale. Therefore, we have tightened them to `rexp(5)`.  

### Model fit
```{r}
m2 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong * congruent +
      b_joint * joint +
      b_sync * sync +
      u_pid[pid] +
      u_dyad[dyad],

    # Fixed effects
    a ~ dnorm(6.1, .2),
    b_cong ~ dnorm(0, .05),
    b_joint ~ dnorm(-.05, .05),
    b_sync ~ dnorm(0, .05),

    # Varying intercepts
    u_pid[pid] ~ dnorm(0, sigma_pid),
    u_dyad[dyad] ~ dnorm(0, sigma_dyad),

    # Scale parameters
    sigma ~ dexp(3),
    sigma_pid ~ dexp(5),
    sigma_dyad ~ dexp(5)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  log_lik = TRUE,
  seed = 42
)
```

```{r}
precis(m2)
```

### Posterior predictive check
```{r}
post2 <- extract.samples(m2)
rt_sim2 <- sim(m2)
rt_ppc2 <- as.vector(rt_sim2)

tibble(
  rt = c(dat$rt, rt_ppc2),
  source = c(
    rep("Observed", length(dat$rt)),
    rep("Posterior predictive", length(rt_ppc2))
  )
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = .5) +
  labs(
    title = "Posterior predictive check — Model 2",
    x = "RT (ms)",
    y = "Density",
    fill = ""
  )
```

### Prior-posterior comparison
```{r}
n_post <- length(post2$a)

prior_post2 <- tibble(
  prior__a = rnorm(n_post, 6.1, .2),
  posterior__a = post2$a,
  prior__cong = rnorm(n_post, 0, .05),
  posterior__cong = post2$b_cong,
  prior__joint = rnorm(n_post, -.05, .05),
  posterior__joint = post2$b_joint,
  prior__sync = rnorm(n_post, 0, .05),
  posterior__sync = post2$b_sync,
  prior__sigma = rexp(n_post, 3),
  posterior__sigma = post2$sigma,
  prior__sigma_pid = rexp(n_post, 5),
  posterior__sigma_pid = post2$sigma_pid,
  prior__sigma_dyad = rexp(n_post, 5),
  posterior__sigma_dyad = post2$sigma_dyad
) |>
  pivot_longer(
    everything(),
    names_to = c("source", "parameter"),
    names_sep = "__"
  )

zoom2 <- prior_post2 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001), xmax = quantile(value, 0.999))

prior_post2 |>
  left_join(zoom2, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = .5) +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison - Model 2",
    x = "",
    y = "Density",
    fill = ""
  )
```

## Model 3 - two-way interactions + varying intercepts
### Prior predictive
```{r}
n_post <- 2000

prior_a <- rnorm(n_post, 6.1, .2)
prior_b_cong <- rnorm(n_post, 0, .05)
prior_b_joint <- rnorm(n_post, -.05, .05)
prior_b_sync <- rnorm(n_post, 0, .05)
prior_b_cong_joint <- rnorm(n_post, 0, .05)
prior_b_cong_sync <- rnorm(n_post, 0, .05)
prior_b_joint_sync <- rnorm(n_post, 0, .05)
prior_sigma <- rexp(n_post, 3)
prior_sigma_pid <- rexp(n_post, 5)
prior_sigma_dyad <- rexp(n_post, 5)

prior_u_pid <- rnorm(n_post, 0, prior_sigma_pid)
prior_u_dyad <- rnorm(n_post, 0, prior_sigma_dyad)

prior_mu <- prior_a +
  prior_b_cong * 1 +
  prior_b_joint * 1 +
  prior_b_sync * 1 +
  prior_b_cong_joint * 1 * 1 +
  prior_b_cong_sync * 1 * 1 +
  prior_b_joint_sync * 1 * 1 +
  prior_u_pid +
  prior_u_dyad

prior_rt <- rlnorm(n_post, meanlog = prior_mu, sdlog = prior_sigma)

tibble(rt = prior_rt) |>
  filter(rt < 5000) |>
  ggplot(aes(x = rt)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check - Model 3",
    subtitle = "Red dashed lines: plausible RT range (150-1800 ms)",
    x = "RT (ms)",
    y = "Density"
  )
```

All three interactions centred at zero so that we presume no interactions unless the data says otherwise. SD of .05 matches the main effects, which is appropriate because interactions should be no larger in magnitude than the main effects themselves.  

### Model fit
```{r}
m3 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong * congruent +
      b_joint * joint +
      b_sync * sync +
      b_cong_joint * congruent * joint +
      b_cong_sync * congruent * sync +
      b_joint_sync * joint * sync +
      u_pid[pid] +
      u_dyad[dyad],

    # Fixed effects
    a ~ dnorm(6.1, .2),
    b_cong ~ dnorm(0, .05),
    b_joint ~ dnorm(-.05, .05),
    b_sync ~ dnorm(0, .05),
    b_cong_joint ~ dnorm(0, .05),
    b_cong_sync ~ dnorm(0, .05),
    b_joint_sync ~ dnorm(0, .05),

    # Varying intercepts
    u_pid[pid] ~ dnorm(0, sigma_pid),
    u_dyad[dyad] ~ dnorm(0, sigma_dyad),

    # Scale parameters
    sigma ~ dexp(3),
    sigma_pid ~ dexp(5),
    sigma_dyad ~ dexp(5)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  log_lik = TRUE,
  seed = 42
)
```

```{r}
precis(m3)
```

### Posterior predictive check
```{r}
post3 <- extract.samples(m3)
rt_sim3 <- sim(m3)
rt_ppc3 <- as.vector(rt_sim3)

tibble(
  rt = c(dat$rt, rt_ppc3),
  source = c(
    rep("Observed", length(dat$rt)),
    rep("Posterior predictive", length(rt_ppc3))
  )
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = .5) +
  labs(
    title = "Posterior predictive check — Model 3",
    x = "RT (ms)",
    y = "Density",
    fill = ""
  )
```

### Prior-posterior comparison
```{r}
post3 <- extract.samples(m3)
n_post <- length(post3$a)

prior_post3 <- tibble(
  prior__a = rnorm(n_post, 6.1, .2),
  posterior__a = post3$a,
  prior__cong = rnorm(n_post, 0, .5),
  posterior__cong = post3$b_cong,
  prior__joint = rnorm(n_post, -.05, .05),
  posterior__joint = post3$b_joint,
  prior__sync = rnorm(n_post, 0, .05),
  posterior__sync = post3$b_sync,
  prior__cong_joint = rnorm(n_post, 0, .05),
  posterior__cong_joint = post3$b_cong_joint,
  prior__cong_sync = rnorm(n_post, 0, .05),
  posterior__cong_sync = post3$b_cong_sync,
  prior__joint_sync = rnorm(n_post, 0, .05),
  posterior__joint_sync = post3$b_joint_sync,
  prior__sigma = rexp(n_post, 3),
  posterior__sigma = post3$sigma,
  prior__sigma_pid = rexp(n_post, 5),
  posterior__sigma_pid = post3$sigma_pid,
  prior__sigma_dyad = rexp(n_post, 5),
  posterior__sigma_dyad = post3$sigma_dyad
) |>
  pivot_longer(
    everything(),
    names_to = c("source", "parameter"),
    names_sep = "__"
  )

zoom3 <- prior_post3 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, .001), xmax = quantile(value, .999))

prior_post3 |>
  left_join(zoom3, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = .5) +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison - Model 3",
    x = "",
    y = "Density",
    fill = ""
  )
```

## Model 4 - three-way interaction
### Prior predictive

Our primary interest: whether synchrony modulates the SSE (the congruency x task-type interaction). Centred at 0 so as to be agnostic about direction. SD of .05 is consistent with all other interaction terms.
```{r}
n_post <- 2000

prior__a <- rnorm(n_post, 6.1, .2)
prior__cong <- rnorm(n_post, 0, .05)
prior__joint <- rnorm(n_post, -.05, .05)
prior__sync <- rnorm(n_post, 0, .05)
prior__cong_joint <- rnorm(n_post, 0, .05)
prior__cong_sync <- rnorm(n_post, 0, .05)
prior__joint_sync <- rnorm(n_post, 0, .05)
prior__cong_joint_sync <- rnorm(n_post, 0, .05)
prior__sigma <- rexp(n_post, 3)
prior__sigma_pid <- rexp(n_post, 5)
prior__sigma_dyad <- rexp(n_post, 5)

prior_u_pid <- rnorm(n_post, 0, prior__sigma_pid)
prior_u_dyad <- rnorm(n_post, 0, prior__sigma_dyad)

prior_mu <- prior__a +
  prior__cong * 1 +
  prior__joint * 1 +
  prior__sync * 1 +
  prior__cong_joint * 1 * 1 +
  prior__cong_sync * 1 * 1 +
  prior__joint_sync * 1 * 1 +
  prior__cong_joint_sync * 1 * 1 * 1 +
  prior_u_pid +
  prior_u_dyad

prior_rt <- rlnorm(n_post, meanlog = prior_mu, sdlog = prior__sigma)

tibble(rt = prior_rt) |>
  filter(rt < 5000) |>
  ggplot(aes(x = rt)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check - Model 4",
    subtitle = "Red dashed lines: plausible RT range (150-1800 ms)",
    x = "RT (ms)",
    y = "Density"
  )
```

### Model fit
```{r}
m4 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong * congruent +
      b_joint * joint +
      b_sync * sync +
      b_cong_joint * congruent * joint +
      b_cong_sync * congruent * sync +
      b_joint_sync * joint * sync +
      b_cong_joint_sync * congruent * joint * sync +
      u_pid[pid] +
      u_dyad[dyad],

    # Fixed effects
    a ~ dnorm(6.1, .2),
    b_cong ~ dnorm(0, .05),
    b_joint ~ dnorm(-.05, .05),
    b_sync ~ dnorm(0, .05),
    b_cong_joint ~ dnorm(0, .05),
    b_cong_sync ~ dnorm(0, .05),
    b_joint_sync ~ dnorm(0, .05),
    b_cong_joint_sync ~ dnorm(0, .05),

    # Varying intercepts
    u_pid[pid] ~ dnorm(0, sigma_pid),
    u_dyad[dyad] ~ dnorm(0, sigma_dyad),

    # Scale parameters
    sigma ~ dexp(3),
    sigma_pid ~ dexp(5),
    sigma_dyad ~ dexp(5)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  log_lik = TRUE,
  seed = 42
)
```

```{r}
m4_precis <- as.data.frame(as.matrix(precis(m4, depth = 1)))
```

```{r}
table3 <- m4_precis |>
  rownames_to_column("Parameter") |>
  mutate(
    Parameter = case_when(
      Parameter == "a" ~ "Intercept",
      Parameter == "b_cong" ~ "Congruency",
      Parameter == "b_joint" ~ "Joint task",
      Parameter == "b_sync" ~ "Synchrony",
      Parameter == "b_cong_joint" ~ "Congruency × Joint",
      Parameter == "b_cong_sync" ~ "Congruency × Synchrony",
      Parameter == "b_joint_sync" ~ "Joint × Synchrony",
      Parameter == "b_cong_joint_sync" ~ "Congruency × Joint × Synchrony",
      Parameter == "sigma" ~ "σ (residual)",
      Parameter == "sigma_pid" ~ "σ (participant)",
      Parameter == "sigma_dyad" ~ "σ (dyad)",
    ),
    across(
      c(mean, sd, `5.5%`, `94.5%`, rhat),
      ~ formatC(.x, format = "f", digits = 2)
    ),
    ess_bulk = formatC(ess_bulk, format = "f", digits = 0)
  ) |>
  rename(
    Mean = mean,
    SD = sd,
    `5.5%` = `5.5%`,
    `94.5%` = `94.5%`,
    Rhat = rhat,
    `ESS (bulk)` = ess_bulk
  ) |>
  kbl(
    caption = "Table 3. Posterior estimates from the Bayesian GLMM (log scale).",
    align = c("l", "r", "r", "r", "r", "r", "r"),
    booktabs = TRUE
  ) |>
  kable_styling(
    latex_options = c("striped", "hold_position"),
    full_width = FALSE
  ) |>
  add_header_above(c(" " = 1, "Posterior" = 4, "Diagnostics" = 2)) |>
  pack_rows("Fixed effects", 1, 8) |>
  pack_rows("Variance components", 9, 11)

table3
```

### Posterior predictive check
```{r}
post4 <- extract.samples(m4)
rt_sim4 <- sim(m4)
rt_ppc4 <- as.vector(rt_sim4)

tibble(
  rt     = c(dat$rt, rt_ppc4),
  source = c(rep("Observed", length(dat$rt)),
             rep("Posterior predictive", length(rt_ppc4)))
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = .5) +
  labs(
    title = "Posterior predictive check — Model 4",
    x = "RT (ms)", y = "Density", fill = ""
  )
```

### Prior-posterior comparison
```{r}
n_post <- length(post4$a)

prior_post4 <- tibble(
  prior__a = rnorm(n_post, 6.1, .2),
  posterior__a = post4$a,
  prior__cong = rnorm(n_post, 0, .05),
  posterior__cong = post4$b_cong,
  prior__joint = rnorm(n_post, -.05, .05),
  posterior__joint = post4$b_joint,
  prior__sync = rnorm(n_post, 0, .05),
  posterior__sync = post4$b_sync,
  prior__cong_joint = rnorm(n_post, 0, .05),
  posterior__cong_joint = post4$b_cong_joint,
  prior__cong_sync = rnorm(n_post, 0, .05),
  posterior__cong_sync = post4$b_cong_sync,
  prior__joint_sync = rnorm(n_post, 0, .05),
  posterior__joint_sync = post4$b_joint_sync,
  prior__cong_joint_sync = rnorm(n_post, 0, .05),
  posterior__cong_joint_sync = post4$b_cong_joint_sync,
  prior__sigma = rexp(n_post, 3),
  posterior__sigma = post4$sigma,
  prior__sigma_pid = rexp(n_post, 5),
  posterior__sigma_pid = post4$sigma_pid,
  prior__sigma_dyad = rexp(n_post, 5),
  posterior__sigma_dyad = post4$sigma_dyad
) |>
  pivot_longer(
    everything(),
    names_to = c("source", "parameter"),
    names_sep = "__"
  )

zoom4 <- prior_post4 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001), xmax = quantile(value, 0.999))

prior_post4 |>
  left_join(zoom4, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = .5) +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison - Model 4",
    x = "",
    y = "Density",
    fill = ""
  )
```

# Model comparison
```{r}
comparison <- compare(m0, m1, m2, m3, m4)

print(comparison)
plot(comparison)
```

`m3` and `m4` not clearly separable.

# Contrasts - SSE by condition
## On log scale
```{r}
# SSE in each condition (on log-RT scale)
SSE_async <- post4$b_cong_joint
SSE_sync <- post4$b_cong_joint + post4$b_cong_joint_sync

# Contrast: difference in SSE between conditions
SSE_contrast <- SSE_sync - SSE_async

# Summarise
contrast_summary <- tibble(
  parameter = c("SSE_async", "SSE_sync", "SSE_contrast"),
  mean = c(mean(SSE_async), mean(SSE_sync), mean(SSE_contrast)),
  sd = c(sd(SSE_async), sd(SSE_sync), sd(SSE_contrast)),
  lower_89 = c(
    quantile(SSE_async, .055),
    quantile(SSE_sync, .055),
    quantile(SSE_contrast, .055)
  ),
  upper_89 = c(
    quantile(SSE_async, .945),
    quantile(SSE_sync, .945),
    quantile(SSE_contrast, .945)
  )
)

contrast_summary

# Plot contrast distribution
tibble(contrast = SSE_contrast) |>
  ggplot(aes(x = contrast)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  labs(
    title = "Posterior contrast: SSE synchrony vs. asynchrony",
    subtitle = "Negative = smaller SSE in synchrony condition",
    x = "SSE_sync - SSE_async (log-RT scale)",
    y = "Density"
  )
```

## Transformed to ms scale
```{r}
# Baseline: reference cell (async, individual, incongruent)
baseline <- exp(post4$a)

# SSE in each condition: exp(a + SSE) - exp(a)
SSE_async_ms <- exp(post4$a + SSE_async) - baseline
SSE_sync_ms <- exp(post4$a + SSE_sync) - baseline
SSE_contrast_ms <- SSE_sync_ms - SSE_async_ms

ms_summary <- tibble(
  parameter = c("SSE_async_ms", "SSE_sync_ms", "SSE_contrast_ms"),
  mean = c(mean(SSE_async_ms), mean(SSE_sync_ms), mean(SSE_contrast_ms)),
  sd = c(sd(SSE_async_ms), sd(SSE_sync_ms), sd(SSE_contrast_ms)),
  lower_89 = c(
    quantile(SSE_async_ms, .055),
    quantile(SSE_sync_ms, .055),
    quantile(SSE_contrast_ms, .055)
  ),
  upper_89 = c(
    quantile(SSE_async_ms, .945),
    quantile(SSE_sync_ms, .945),
    quantile(SSE_contrast_ms, .945)
  )
)

ms_summary

# Plot contrast in ms
tibble(contrast_ms = SSE_contrast_ms) |>
  ggplot(aes(x = contrast_ms)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  labs(
    title = "Figure 1. Posterior contrast: SSE synchrony vs. asynchrony (ms)",
    subtitle = "Positive values indicate a larger SSE in synchrony condition.\nNegative values indicate a larger SSE in the asynchrony condition.",
    x = "SSE sync - SSE async (ms)",
    y = "Density"
  )


dens <- density(SSE_contrast_ms)
dens_df <- tibble(x = dens$x, y = dens$y) |>
mutate(region = ifelse(x > 0, "positive", "negative"))

dens_df |>
  ggplot(aes(x = x, y = y, fill = region)) +
  geom_area(alpha = .5) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
scale_fill_manual(
  values = c("positive" = color1, "negative" = color2),
  labels = c("positive" = "SSE larger in synchrony", "negative" = "SSE larger in asynchrony")
  ) +
  labs(
    title = "Figure 1. Posterior contrast: SSE synchrony vs. asynchrony (ms)",
    # subtitle = "Positive values indicate a larger SSE in synchrony condition.\nNegative values indicate a larger SSE in the asynchrony condition.",
    x = "SSE sync - SSE async (ms)",
    y = "Density",
    fill = NULL
  ) +
  theme(legend.position = "bottom")
```

```{r}
mean(SSE_contrast_ms < 0)
```

# Sensitivity check
We widen priors to see if this affects posteriors.
```{r}
m4_wide <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong * congruent +
      b_joint * joint +
      b_sync * sync +
      b_cong_joint * congruent * joint +
      b_cong_sync * congruent * sync +
      b_joint_sync * joint * sync +
      b_cong_joint_sync * congruent * joint * sync +
      u_pid[pid] +
      u_dyad[dyad],

    # Fixed effects — wider priors
    a ~ dnorm(6.1, .5), # was .2
    b_cong ~ dnorm(0, .2), # was .05
    b_joint ~ dnorm(-.05, .2), # was .05
    b_sync ~ dnorm(0, .2), # was .05
    b_cong_joint ~ dnorm(0, .2), # was .05
    b_cong_sync ~ dnorm(0, .2), # was .05
    b_joint_sync ~ dnorm(0, .2), # was .05
    b_cong_joint_sync ~ dnorm(0, .2), # was .05

    # Varying intercepts
    u_pid[pid] ~ dnorm(0, sigma_pid),
    u_dyad[dyad] ~ dnorm(0, sigma_dyad),

    # Scale parameters — wider
    sigma ~ dexp(1), # was 3
    sigma_pid ~ dexp(1), # was 5
    sigma_dyad ~ dexp(1) # was 5
  ),
  data = dat,
  chains = 4,
  cores = 4,
  log_lik = TRUE,
  seed = 42
)
```

```{r}
m4_wide_precis <- as.data.frame(as.matrix(precis(m4_wide, depth = 1)))
```

```{r}
table3_wide <- m4_wide_precis |>
  rownames_to_column("Parameter") |>
  mutate(
    Parameter = case_when(
      Parameter == "a" ~ "Intercept",
      Parameter == "b_cong" ~ "Congruency",
      Parameter == "b_joint" ~ "Joint task",
      Parameter == "b_sync" ~ "Synchrony",
      Parameter == "b_cong_joint" ~ "Congruency × Joint",
      Parameter == "b_cong_sync" ~ "Congruency × Synchrony",
      Parameter == "b_joint_sync" ~ "Joint × Synchrony",
      Parameter == "b_cong_joint_sync" ~ "Congruency × Joint × Synchrony",
      Parameter == "sigma" ~ "σ (residual)",
      Parameter == "sigma_pid" ~ "σ (participant)",
      Parameter == "sigma_dyad" ~ "σ (dyad)",
    ),
    across(
      c(mean, sd, `5.5%`, `94.5%`, rhat),
      ~ formatC(.x, format = "f", digits = 2)
    ),
    ess_bulk = formatC(ess_bulk, format = "f", digits = 0)
  ) |>
  rename(
    Mean = mean,
    SD = sd,
    `5.5%` = `5.5%`,
    `94.5%` = `94.5%`,
    Rhat = rhat,
    `ESS (bulk)` = ess_bulk
  ) |>
  kbl(
    caption = "SENSITIVITY Table 3. Posterior estimates from the Bayesian GLMM (log scale).",
    align = c("l", "r", "r", "r", "r", "r", "r"),
    booktabs = TRUE
  ) |>
  kable_styling(
    latex_options = c("striped", "hold_position"),
    full_width = FALSE
  ) |>
  add_header_above(c(" " = 1, "Posterior" = 4, "Diagnostics" = 2)) |>
  pack_rows("Fixed effects", 1, 8) |>
  pack_rows("Variance components", 9, 11)

table3_wide
```

## Posterior predictive check
```{r}
post4_wide <- extract.samples(m4_wide)
rt_sim4_wide <- sim(m4_wide)
rt_ppc4_wide <- as.vector(rt_sim4_wide)

tibble(
  rt     = c(dat$rt, rt_ppc4_wide),
  source = c(rep("Observed", length(dat$rt)),
             rep("Posterior predictive", length(rt_ppc4_wide)))
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = .5) +
  labs(
    title = "Posterior predictive check — Model 4",
    x = "RT (ms)", y = "Density", fill = ""
  )
```

## Prior-posterior comparison
```{r}
n_post <- length(post4_wide$a)

prior_post4_wide <- tibble(
  prior__a = rnorm(n_post, 6.1, .5),
  posterior__a = post4_wide$a,
  prior__cong = rnorm(n_post, 0, .2),
  posterior__cong = post4_wide$b_cong,
  prior__joint = rnorm(n_post, -.05, .2),
  posterior__joint = post4_wide$b_joint,
  prior__sync = rnorm(n_post, 0, .2),
  posterior__sync = post4_wide$b_sync,
  prior__cong_joint = rnorm(n_post, 0, .2),
  posterior__cong_joint = post4_wide$b_cong_joint,
  prior__cong_sync = rnorm(n_post, 0, .2),
  posterior__cong_sync = post4_wide$b_cong_sync,
  prior__joint_sync = rnorm(n_post, 0, .2),
  posterior__joint_sync = post4_wide$b_joint_sync,
  prior__cong_joint_sync = rnorm(n_post, 0, .2),
  posterior__cong_joint_sync = post4_wide$b_cong_joint_sync,
  prior__sigma = rexp(n_post, 1),
  posterior__sigma = post4_wide$sigma,
  prior__sigma_pid = rexp(n_post, 1),
  posterior__sigma_pid = post4_wide$sigma_pid,
  prior__sigma_dyad = rexp(n_post, 1),
  posterior__sigma_dyad = post4_wide$sigma_dyad
) |>
  pivot_longer(
    everything(),
    names_to = c("source", "parameter"),
    names_sep = "__"
  )

zoom4_wide <- prior_post4_wide |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, .001), xmax = quantile(value, .999))

prior_post4_wide |>
  left_join(zoom4_wide, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = .5) +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "SENSITIVITY Prior-posterior comparison - Model 4",
    x = "",
    y = "Density",
    fill = ""
  )
```

## Contrasts - SSE by condition
### On log scale
```{r}
# SSE in each condition (on log-RT scale)
SSE_async <- post4_wide$b_cong_joint
SSE_sync <- post4_wide$b_cong_joint + post4_wide$b_cong_joint_sync

# Contrast: difference in SSE between conditions
SSE_contrast <- SSE_sync - SSE_async

# Summarise
contrast_summary <- tibble(
  parameter = c("SSE_async", "SSE_sync", "SSE_contrast"),
  mean = c(mean(SSE_async), mean(SSE_sync), mean(SSE_contrast)),
  sd = c(sd(SSE_async), sd(SSE_sync), sd(SSE_contrast)),
  lower_89 = c(
    quantile(SSE_async, .055),
    quantile(SSE_sync, .055),
    quantile(SSE_contrast, .055)
  ),
  upper_89 = c(
    quantile(SSE_async, .945),
    quantile(SSE_sync, .945),
    quantile(SSE_contrast, .945)
  )
)

contrast_summary

# Plot contrast distribution
tibble(contrast = SSE_contrast) |>
  ggplot(aes(x = contrast)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  labs(
    title = "SENSITIVITY Posterior contrast: SSE synchrony vs. asynchrony",
    subtitle = "Negative = smaller SSE in synchrony condition",
    x = "SSE_sync - SSE_async (log-RT scale)",
    y = "Density"
  )
```

### Transformed to ms scale
```{r}
# Baseline: reference cell (async, individual, incongruent)
baseline <- exp(post4_wide$a)

# SSE in each condition: exp(a + SSE) - exp(a)
SSE_async_ms <- exp(post4_wide$a + SSE_async) - baseline
SSE_sync_ms <- exp(post4_wide$a + SSE_sync) - baseline
SSE_contrast_ms <- SSE_sync_ms - SSE_async_ms

ms_summary <- tibble(
  parameter = c("SSE_async_ms", "SSE_sync_ms", "SSE_contrast_ms"),
  mean = c(mean(SSE_async_ms), mean(SSE_sync_ms), mean(SSE_contrast_ms)),
  sd = c(sd(SSE_async_ms), sd(SSE_sync_ms), sd(SSE_contrast_ms)),
  lower_89 = c(
    quantile(SSE_async_ms, .055),
    quantile(SSE_sync_ms, .055),
    quantile(SSE_contrast_ms, .055)
  ),
  upper_89 = c(
    quantile(SSE_async_ms, .945),
    quantile(SSE_sync_ms, .945),
    quantile(SSE_contrast_ms, .945)
  )
)

ms_summary

# Plot contrast in ms
tibble(contrast_ms = SSE_contrast_ms) |>
  ggplot(aes(x = contrast_ms)) +
  geom_density(fill = color1, alpha = .5) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  labs(
    title = "SENSITIVITY Figure 1. Posterior contrast: SSE synchrony vs. asynchrony (ms)",
    subtitle = "Positive values indicate a larger SSE in synchrony condition.\nNegative values indicate a larger SSE in the asynchrony condition.",
    x = "SSE sync - SSE async (ms)",
    y = "Density"
  )

dens <- density(SSE_contrast_ms)
dens_df <- tibble(x = dens$x, y = dens$y) |>
mutate(region = ifelse(x > 0, "positive", "negative"))

dens_df |>
  ggplot(aes(x = x, y = y, fill = region)) +
  geom_area(alpha = .5) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
scale_fill_manual(
  values = c("positive" = color1, "negative" = color2),
  labels = c("positive" = "SSE larger in synchrony", "negative" = "SSE larger in asynchrony")
  ) +
  labs(
    title = "SENSITIVITY Figure 1. Posterior contrast: SSE synchrony vs. asynchrony (ms)",
    # subtitle = "Positive values indicate a larger SSE in synchrony condition.\nNegative values indicate a larger SSE in the asynchrony condition.",
    x = "SSE sync - SSE async (ms)",
    y = "Density",
    fill = NULL
  ) +
  theme(legend.position = "bottom")
```

```{r}
mean(SSE_contrast_ms < 0)
```

# Parameter recovery check
Since `m4` is our main model, it makes sense to check whether it is even able to pick up true signals in the data. Therefore, we simulate data with true effects to see if the model recovers these.

```{r}
# true parameter values
true <- list(
  a = 6.1,
  b_cong = -.04,
  b_joint = -.05,
  b_sync = 0,
  b_cong_joint = -.03,
  b_cong_sync = 0,
  b_joint_sync = 0,
  b_cong_joint_sync = .1, # also tried .03 and 0
  sigma = .33,
  sigma_pid = .20,
  sigma_dyad = .20
)

# simulate data
set.seed(42)

n_dyads <- 9
n_pid <- n_dyads * 2 # 18 participants
n_trials_individual <- 60 # per participant
n_trials_joint <- 60 # per dyad (logged twice, once per participant)

# Varying intercepts
u_dyad <- rnorm(n_dyads, 0, true$sigma_dyad)
u_pid <- rnorm(n_pid, 0, true$sigma_pid)

# Build trial-level data frame
sim_data <- expand_grid(
  dyad_id = 1:n_dyads,
  congruent = c(0, 1), # 0 = incongruent, 1 = congruent
  joint = c(0, 1), # 0 = individual, 1 = joint
  sync = c(0, 1) # 0 = async, 1 = sync
) |>
  # Each combination gets n_trials_individual/4 trials per participant
  # (balanced: 2 congruency × 2 task × 2 sync levels --> 60/4 = 15 trials each)
  slice(rep(1:n(), each = 15)) |>
  mutate(
    pid = if_else(
      joint == 0,
      (dyad_id - 1) * 2 + sample(1:2, n(), replace = TRUE),
      (dyad_id - 1) * 2 + 1
    ), # joint: attribute to P1 for simplicity
    mu = true$a +
      true$b_cong * congruent +
      true$b_joint * joint +
      true$b_sync * sync +
      true$b_cong_joint * congruent * joint +
      true$b_cong_sync * congruent * sync +
      true$b_joint_sync * joint * sync +
      true$b_cong_joint_sync * congruent * joint * sync +
      u_pid[pid] +
      u_dyad[dyad_id],
    rt = rlnorm(n(), meanlog = mu, sdlog = true$sigma)
  )

# fit model
m4_sim <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong * congruent +
      b_joint * joint +
      b_sync * sync +
      b_cong_joint * congruent * joint +
      b_cong_sync * congruent * sync +
      b_joint_sync * joint * sync +
      b_cong_joint_sync * congruent * joint * sync +
      u_pid[pid] +
      u_dyad[dyad_id],
    # Fixed effects
    a ~ dnorm(6.1, .2),
    b_cong ~ dnorm(0, .05),
    b_joint ~ dnorm(-.05, .05),
    b_sync ~ dnorm(0, .05),
    b_cong_joint ~ dnorm(0, .05),
    b_cong_sync ~ dnorm(0, .05),
    b_joint_sync ~ dnorm(0, .05),
    b_cong_joint_sync ~ dnorm(0, .05),
    # Residual
    sigma ~ dexp(3),
    # Random intercepts
    u_pid[pid] ~ dnorm(0, sigma_pid),
    u_dyad[dyad_id] ~ dnorm(0, sigma_dyad),
    sigma_pid ~ dexp(5),
    sigma_dyad ~ dexp(5)
  ),
  data = sim_data |>
    select(rt, congruent, joint, sync, pid, dyad_id) |>
    as.list(),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 500,
  seed = 42
)

# recovery check
post <- extract.samples(m4_sim)

focal_params <- c(
  "a",
  "b_cong",
  "b_joint",
  "b_sync",
  "b_cong_joint",
  "b_cong_sync",
  "b_joint_sync",
  "b_cong_joint_sync"
)

recovery <- map_dfr(focal_params, function(p) {
  samples <- post[[p]]
  tibble(
    param = p,
    true_val = true[[p]],
    mean = mean(samples),
    lower89 = quantile(samples, .055),
    upper89 = quantile(samples, .945),
    recovered = true[[p]] >= quantile(samples, .055) &
      true[[p]] <= quantile(samples, .945)
  )
})

print(recovery)

# Plot
recovery |>
  ggplot(aes(x = param, y = mean, ymin = lower89, ymax = upper89)) +
  geom_pointrange() +
  geom_point(aes(y = true_val), colour = "red", size = 3, shape = 4) +
  coord_flip() +
  labs(
    title = "Parameter recovery — Model 4",
    subtitle = "Black: posterior mean + 89 % CI. Red cross: true value.",
    x = NULL,
    y = "Estimate (log scale)"
  )
```

Model struggles to pick up true effects at n_dyad = 9.