---
title: "Analysis Script for Social and Cultural Dynamics 2026"
author: "Andrea Wiwe Kjær-Hansen, Mikkel Krøjer Svendsen"
format: html
---

# Data and R setup

```{r, message=FALSE}
# Packages
library(rethinking)
# library(dagitty)
# library(ggdag)
library(tidyverse)
library(lmerTest)

# Theme settings
theme_set(theme_minimal())

# data directory
source("data_dir.R")
```

```{r}
# read files helper
read_files <- function(pattern) {
  list.files(
    data_dir,
    pattern = pattern,
    full.names = TRUE
  ) |>
    map_dfr(~ read.delim2(.x, sep = ","))
}
```

```{r}
# read files
d_simon <- read_files("simon") |>
  mutate(
    rt_ms = as.numeric(ifelse(rt_ms == "", NA, rt_ms)),
    response_key = ifelse(response_key == "", NA, response_key),
    side_congruent = (shape == "square" & side == "left") |
      (shape == "diamond" & side == "right")
  )

d_tapping <- read_files("tapping")

d_info <- read_files("group_info") |> 
  mutate(
    gender = ifelse(dyad_id == 4 & participant == 2, "Male", gender)
  )
```

# Descriptive statistics and data cleaning
## Participants
```{r}
age <- d_info |>
  summarize(
    age_mean = mean(age),
    age_sd = sd(age)
  )

gender <- d_info |>
  count(gender) |>
  mutate(
    prop = n / sum(n)
  )

condition <- d_tapping |>
  distinct(dyad_id, .keep_all = T) |>
  count(condition, name = "n_dyads")

print(age)
print(gender)
print(condition)
```

## Calculating the Social Simon Effect (SSE)

What is the SSE? In the joint go/no-go Simon task, each participant only responds to one shape. 

The Simon Effect shouldn't exist individually. BUT it often does appear as if participants mentally represent their partner's role. 

So, the SSE: Faster/more accurate responses when the stimulus appears on your response-congruent side, even though congruence is only meaningful relative to the dyad.

$$\text{SSE} = \text{RT}_{\text{congruent}} − \text{RT}_{\text{incongruent}}$$

First, add condition column to data

```{r}
# add condition to d_simon
d_simon <- d_simon |>
  left_join(d_tapping |> distinct(dyad_id, condition), by = "dyad_id")

str(d_simon)
```

We only look at joint trials that are correct go trials (`responded == 1` and `correct == 1`)

```{r}
d_joint <- d_simon |>
  filter(task == "joint" & responded == 1 & correct == 1) |>
  mutate(participant_id = paste(dyad_id, participant, sep = "_"))

str(d_joint)
```

We can condition on condition and side congruence:
```{r}
d_joint |>
  count(participant_id, side, side_congruent)
```

We can also collapse to one SSE value per dyad (throwing away a lot of information in the process...)

```{r}
dyad_sse <- d_joint |>
  group_by(dyad_id, condition, side_congruent) |>
  summarise(mean_rt = mean(rt_ms), .groups = "drop") |>
  pivot_wider(names_from = side_congruent, values_from = mean_rt) |>
  mutate(SSE = `FALSE` - `TRUE`)

dyad_sse
```

Summarizing condition and side congruence
```{r}
d_joint |>
  group_by(condition, side_congruent) |>
  summarise(
    mean_rt = mean(rt_ms),
    sd_rt = sd(rt_ms),
    n = n(),
    .groups = "drop"
  )
```

Summarizing condition
```{r}
d_joint |>
  group_by(condition, side_congruent) |>
  summarise(mean_rt = mean(rt_ms), .groups = "drop") |>
  pivot_wider(names_from = side_congruent, values_from = mean_rt) |>
  mutate(SSE = `FALSE` - `TRUE`)
```

Huh, synchrony dyads have lower SSE than asynchrony dyads. This seems to go against theoretical expectation...

## Visualizations
RT distributions (overlaid)
```{r}
d_joint |>
  mutate(
    side_congruent = ifelse(side_congruent, "Congruent", "Incongruent"),
    condition = str_to_title(condition)
  ) |>
  ggplot(aes(x = rt_ms, fill = side_congruent)) +
  geom_density(alpha = 0.5) +
  facet_wrap(vars(condition)) +
  labs(
    x = "Reaction Time (ms)",
    y = "Density",
    fill = "Congruence",
    title = "RT Distributions by Condition and Congruence"
  )
```
RT distributions (faceted)
```{r}
d_joint |>
  mutate(
    side_congruent = ifelse(side_congruent, "Congruent", "Incongruent"),
    condition = str_to_title(condition)
  ) |>
  ggplot(aes(x = rt_ms, fill = side_congruent)) +
  geom_density(alpha = 0.5) +
  facet_wrap(vars(side_congruent, condition)) +
  labs(
    x = "Reaction Time (ms)",
    y = "Density",
    fill = "Congruence",
    title = "RT Distributions by Condition and Congruence"
  )
```

SSE distributions (condition mean and individual dyads)
```{r}
dyad_sse |>
  mutate(condition = str_to_title(condition)) |>
  ggplot(aes(x = condition, y = SSE, fill = condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_boxplot(alpha = 0.5) +
  geom_jitter(width = 0.05, size = 3, alpha = .5) +
  labs(
    x = "Condition",
    y = "Social Simon Effect (ms)\n[Incongruent − Congruent RT]",
    title = "SSE per Dyad by Condition"
  ) +
  theme(legend.position = "none")
```

Congruence effect per dyad by condition
```{r}
d_joint |>
  group_by(dyad_id, condition, side_congruent) |>
  summarise(mean_rt = mean(rt_ms), .groups = "drop") |>
  mutate(
    side_congruent = ifelse(side_congruent, "Congruent", "Incongruent"),
    condition = str_to_title(condition)
  ) |>
  ggplot(aes(
    x = side_congruent,
    y = mean_rt,
    group = dyad_id,
    colour = factor(dyad_id)
  )) +
  geom_line() +
  geom_point(size = 3) +
  facet_wrap(~condition) +
  labs(
    x = "",
    y = "Mean RT (ms)",
    colour = "Dyad",
    title = "Congruence Effect per Dyad by Condition"
  )
```

# Analysis
## Simple t-test
First, we can do a t-test. Since this is a between-dyad experiment, we assume the two conditions to be independent. We use Welch's test as it does not assume equal variances across groups.
```{r}
t.test(SSE ~ condition, data = dyad_sse)
```

The difference between conditions is trivially small. But it's worth noting that it goes against our expectations: Asynchrony condition dyads show a larger SSE than do the synchrony condition dyads.

## Multilevel model
Beefing up our analysis, we can do a multilevel model to account for variances at different levels of our experiment.  

First, we fit this model:
- Outcome: Trial-level RT  
- Fixed effects: Congruence, condition, and their interaction  
- Random effects: Dyad and participant  
```{r}
m1 <- lmer(
  rt_ms ~ side_congruent * condition + (1 | dyad_id) + (1 | participant_id),
  data = d_joint
)

summary(m1)
```

The fit is singular, i.e., the model cannot distinguish between dyad-level and participant-level variances. This makes sense since we only have two participants in each dyad, and they always have the same role (pressing either the `Z` or `M` key).  

We try a new model using only dyad-level variances as random effects.

```{r}
m2 <- lmer(
  rt_ms ~ side_congruent * condition + (1 | dyad_id),
  data = d_joint
)

summary(m2)
```

## Bayesian models
### Model 1
#### Data
We mirror the frequentist `m2` in structure. First, we define the data list.
```{r}
dat <- d_joint |>
  mutate(
    cong = as.integer(side_congruent),
    cond = as.integer(condition == "synchrony"),
    dyad = as.integer(factor(dyad_id)),
    rt = as.numeric(rt_ms)
  ) |>
  select(rt, cong, cond, dyad) |>
  as.list()

dat$N_dyads <- max(dat$dyad)
```

#### Prior predictive

```{r}
mb1_prior <- ulam(
  alist(
    rt ~ dnorm(mu, sigma),
    mu <- a +
          b_cong * cong +
          b_cond * cond +
          b_interaction * cong * cond +
          a_dyad[dyad],

    a ~ dnorm(400, 50), # plausible mean RT in ms
    b_cong ~ dnorm(0, 20), # effects in ms; 20ms SD is weakly informative
    b_cond ~ dnorm(0, 20),
    b_interaction ~ dnorm(0, 20),

    a_dyad[dyad] ~ dnorm(0, sigma_dyad),
    sigma_dyad ~ dexp(1),
    sigma ~ dexp(1)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  sample_prior = TRUE
)

prior <- extract.prior(mb1_prior, n = 10000)
saveRDS(mb1_prior, file = "model_and_samples/mb1_prior_model.rds")
saveRDS(prior, file = "model_and_samples/mb1_prior.rds")
```


```{r}
mb1_prior <- readRDS("model_and_samples/mb1_prior_model.rds")
prior <- readRDS("model_and_samples/mb1_prior.rds")
```

We do a prior predictive check:
```{r}
# SSE per condition
sse_async_prior <- (prior$a + prior$b_cong) - prior$a
sse_sync_prior <- (prior$a +
  prior$b_cond +
  prior$b_cong +
  prior$b_interaction) -
  (prior$a + prior$b_cond)
delta_prior <- sse_sync_prior - sse_async_prior # simplifies to just b_interaction

# Plot implied RT distributions
par(mfrow = c(2, 2))
dens(prior$a, main = "Incongruent / Async", xlab = "RT (ms)")
dens(prior$a + prior$b_cong, main = "Congruent / Async", xlab = "RT (ms)")
dens(prior$a + prior$b_cond, main = "Incongruent / Sync", xlab = "RT (ms)")
dens(
  prior$a + prior$b_cond + prior$b_cong + prior$b_interaction,
  main = "Congruent / Sync",
  xlab = "RT (ms)"
)
par(mfrow = c(1, 1))

# Prior on SSE difference
data.frame(value = delta_prior) |>
  ggplot(aes(x = value)) +
  geom_density(fill = "forestgreen", colour = "forestgreen", alpha = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  labs(
    x = "SSE difference: Sync - Async (ms)",
    y = "Density",
    title = "Prior: SSE Difference (ms scale)"
  ) +
  theme_minimal()
```

#### Model fit

```{r}
mb1 <- ulam(
  alist(
    rt ~ dnorm(mu, sigma),
    mu <- a +
          b_cong        * cong +
          b_cond        * cond +
          b_interaction * cong * cond +
          a_dyad[dyad],

    a             ~ dnorm(400, 50),
    b_cong        ~ dnorm(0, 20),
    b_cond        ~ dnorm(0, 20),
    b_interaction ~ dnorm(0, 20),

    a_dyad[dyad]  ~ dnorm(0, sigma_dyad),
    sigma_dyad    ~ dexp(1),
    sigma         ~ dexp(1)
  ),
  data    = dat,
  chains  = 4,
  cores   = 4,
  iter    = 2000,
  log_lik = TRUE
)

post <- extract.samples(mb1, n = 10000)
saveRDS(mb1,  file = "model_and_samples/mb1.rds")
saveRDS(post, file = "model_and_samples/mb1_post.rds")
```

```{r}
mb1 <- readRDS("model_and_samples/mb1.rds")
post <- readRDS("model_and_samples/mb1_post.rds")
```

First, we're interested in `b_interaction`. We predict it to be negative (i.e., congruent trials are faster, and this advantage is strongest in synchrony dyads).

```{r}
precis(mb1)
```

The posterior of the interaction term (b_interaction) was centred near zero (mean = 2.96, 89 % CI: [-9.96, 16.52]). This indicates no credible evidence that the Social Simon Effect differed between synchrony and asynchrony conditions.

```{r}
# b_interaction on ms scale
data.frame(
  value        = c(prior$b_interaction, post$b_interaction),
  Distribution = c(rep("Prior",     length(prior$b_interaction)),
                   rep("Posterior", length(post$b_interaction)))
) |>
  ggplot(aes(x = value, fill = Distribution, colour = Distribution)) +
  geom_density(alpha = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_fill_manual(values   = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  scale_colour_manual(values = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  labs(x = "b_interaction (ms)", y = "Density",
       title = "Prior vs Posterior: Interaction Effect") +
  theme_minimal()

# SSE difference on ms scale
delta_post <- post$b_interaction   # for Normal model this is already in ms

data.frame(
  value        = c(delta_prior, delta_post),
  Distribution = c(rep("Prior",     length(delta_prior)),
                   rep("Posterior", length(delta_post)))
) |>
  ggplot(aes(x = value, fill = Distribution, colour = Distribution)) +
  geom_density(alpha = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_fill_manual(values   = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  scale_colour_manual(values = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  labs(x = "SSE difference: Sync - Async (ms)", y = "Density",
       title = "Prior vs Posterior: SSE Difference (ms scale)") +
  theme_minimal()
```
The posterior is substantially narrower than the prior which suggests that the data was informative but did not shift meaningfully away from zero in either direction.

So what have we learned? Well, we weren't sure which way the effect was going to go (or even if it was going to go at all). It seems like it isn't going to go at all. And if it is, we still don't know which way. This is a fun exam.

Revision notes that I forgot to add lol:
- Don't assume normal likelihood for RTs, silly. Use log-normal or exGaussian (EMG).
- Change random effects structure: Varying slopes should be modeled whenever the treatment effect is expected to vary across clusters (should be allowed to correlate with intercepts via an LKJ-prior covariance matrix).
- A more complete model would include a participant-level random effect nested inside dyads.
- Do prior predictive simulations (also use as power analysis)!

### Model 2 - lognorm likelihood

Now we do NOT standardize RTs. Instead, we use a log_normal likelihood.

#### Data
```{r}
dat <- d_joint |>
  mutate(
    cong = as.integer(side_congruent),
    cond = as.integer(condition == "synchrony"),
    dyad = as.integer(factor(dyad_id)),
    rt = as.numeric(rt_ms) # raw ms, no z-scoring
  ) |>
  select(rt, cong, cond, dyad) |>
  as.list()

dat$N_dyads <- max(dat$dyad)
```

#### Prior predictive
```{r}
mb2_prior <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong * cong +
      b_cond * cond +
      b_interaction * cong * cond +
      a_dyad[dyad],

    # Intercept: exp(6) ≈ 403ms, SD of 0.3 keeps prior mass
    # roughly between exp(5.4)≈220ms and exp(6.6)≈735ms
    a ~ dnorm(6, 0.3),

    # Slopes on log scale: 0.1 ≈ 10% shift in RT
    # SD of 0.2 allows effects up to ~20% while remaining sceptical
    b_cong ~ dnorm(0, 0.2),
    b_cond ~ dnorm(0, 0.2),
    b_interaction ~ dnorm(0, 0.2),

    a_dyad[dyad] ~ dnorm(0, sigma_dyad),
    sigma_dyad ~ dexp(5), # tighter: 8 dyads can't estimate
    # large between-dyad variance reliably
    sigma ~ dexp(1)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  sample_prior = TRUE
)

prior <- extract.prior(mb2_prior, n = 10000)
saveRDS(mb2_prior, file = "model_and_samples/mb2_prior_model.rds")
saveRDS(prior, file = "model_and_samples/mb2_prior.rds")
```

```{r}
mb2_prior <- readRDS("model_and_samples/mb2_prior_model.rds")
prior <- readRDS("model_and_samples/mb2_prior.rds")
```

We do a prior predictive check:
```{r}
# Extract prior predictive RTs for each condition combination
# cong=0,cond=0: incongruent, asynchrony (baseline)
# cong=1,cond=0: congruent, asynchrony
# cong=0,cond=1: incongruent, synchrony
# cong=1,cond=1: congruent, synchrony

n <- length(prior$a)

rt_ic_as <- exp(
  prior$a
)
rt_c_as <- exp(
  prior$a + prior$b_cong
)
rt_ic_s <- exp(
  prior$a + prior$b_cond
)
rt_c_s <- exp(
  prior$a + prior$b_cong + prior$b_cond + prior$b_interaction
)

# Plot
par(mfrow = c(2, 2))
dens(
  rt_ic_as,
  main = "Incongruent / Async",
  xlab = "RT (ms)",
  xlim = c(0, 2000)
)
dens(rt_c_as, main = "Congruent / Async", xlab = "RT (ms)", xlim = c(0, 2000))
dens(rt_ic_s, main = "Incongruent / Sync", xlab = "RT (ms)", xlim = c(0, 2000))
dens(rt_c_s, main = "Congruent / Sync", xlab = "RT (ms)", xlim = c(0, 2000))
#par(mfrow = c(1, 1))
```

```{r}
# Simon effect = congruent RT minus incongruent RT, within each condition
# On the log scale this is just b_cong (async) or b_cong + b_interaction (sync)
# But we want it on the ms scale, so compute as a ratio or raw difference

# Median RT per cell (ignoring dyad variability for clarity)
rt_ic_as <- exp(prior$a)
rt_c_as  <- exp(prior$a + prior$b_cong)
rt_ic_s  <- exp(prior$a + prior$b_cond)
rt_c_s   <- exp(prior$a + prior$b_cond + prior$b_cong + prior$b_interaction)

# Simon effect per condition (congruent - incongruent, in ms)
sse_async <- rt_c_as - rt_ic_as
sse_sync  <- rt_c_s  - rt_ic_s

# The interaction: difference in SSE between conditions
delta_sse <- sse_sync - sse_async

dens(delta_sse, main = "Prior: SSE difference (Sync - Async)", xlab = "ms")
abline(v = 0, lty = 2, col = "red")
```

Notice that the prior straddles 0. This is fine, as we find both positive and negative effects plausible a priori.

#### Model fit
```{r}
mb2 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
          b_cong        * cong +
          b_cond        * cond +
          b_interaction * cong * cond +
          a_dyad[dyad],

    a             ~ dnorm(6, 0.3),
    b_cong        ~ dnorm(0, 0.2),
    b_cond        ~ dnorm(0, 0.2),
    b_interaction ~ dnorm(0, 0.2),

    a_dyad[dyad]  ~ dnorm(0, sigma_dyad),
    sigma_dyad    ~ dexp(5),
    sigma         ~ dexp(1)
  ),
  data    = dat,
  chains  = 4,
  cores   = 4,
  iter    = 2000,
  log_lik = TRUE
)

post <- extract.samples(mb2, n = 10000)
saveRDS(mb2,  file = "model_and_samples/mb2.rds")
saveRDS(post, file = "model_and_samples/mb2_post.rds")
```

```{r}
mb2 <- readRDS("model_and_samples/mb2.rds")
post <- read_rds("model_and_samples/mb2_post.rds")
```

```{r}
precis(mb2, 2)
```

```{r}
# Prior vs posterior for b_interaction
dens(prior$b_interaction, lty = 2, 
     xlab = "b_interaction", main = "Prior vs Posterior: interaction", ylim=c(0,18), xlim=c(-.5,.5))
dens(post$b_interaction, add = TRUE)
abline(v = 0, lty = 3, col = "red")
legend("topright", legend = c("prior", "posterior"), lty = c(2, 1))
```

```{r}
rt_ic_as <- exp(post$a)
rt_c_as  <- exp(post$a + post$b_cong)
rt_ic_s  <- exp(post$a + post$b_cond)
rt_c_s   <- exp(post$a + post$b_cond + post$b_cong + post$b_interaction)

sse_async <- rt_c_as - rt_ic_as
sse_sync  <- rt_c_s  - rt_ic_s
delta_sse <- sse_sync - sse_async

dens(delta_sse, main = "Posterior: SSE difference (Sync - Async)", xlab = "ms")
abline(v = 0, lty = 2, col = "red")
PI(delta_sse)       # 89% interval
mean(delta_sse)     # posterior mean in ms
```

In ggplot:
```{r}
# Image 1: b_interaction prior vs posterior
data.frame(
  value        = c(prior$b_interaction, post$b_interaction),
  Distribution = c(rep("Prior",     length(prior$b_interaction)),
                   rep("Posterior", length(post$b_interaction)))
) |>
  ggplot(aes(x = value, fill = Distribution, colour = Distribution)) +
  geom_density(alpha = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_fill_manual(values   = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  scale_colour_manual(values = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  labs(x = "b_interaction", y = "Density",
       title = "Prior vs Posterior: Interaction Effect") +
  theme_minimal()

# Image 2: SSE difference on ms scale (posterior only)
data.frame(value = delta_post) |>
  ggplot(aes(x = value)) +
  geom_density(fill = "steelblue", colour = "steelblue", alpha = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  labs(x = "ms", y = "Density",
       title = "Posterior: SSE Difference (Sync - Async)") +
  theme_minimal()
```

```{r}
plot(compare(mb1, mb2))
```

### Model 3 - varying slopes on congruence
Now we add varying slopes on the congruence variable at the dyad level. We correlate this with the dyad intercept via an LKJ covariance matrix. So, each dyad gets its own (intercept, congruence slope)-pair drawn from a bivariate normal.

#### Data
Now, we add participant_id to `dat` so that we can check whether it makes sense to do participant-level intercepts.  

Note: `b_cond` and `b_interaction` remain fixed effects because condition is a between-dyads variable. There's no within-dyad variation in condition from which to estimate a varying slope.  
```{r}
dat <- d_joint |>
  mutate(
    cong        = as.integer(side_congruent),
    cond        = as.integer(condition == "synchrony"),
    dyad        = as.integer(factor(dyad_id)),
    participant = as.integer(factor(participant_id)),
    rt          = as.numeric(rt_ms)
  ) |>
  select(rt, cong, cond, dyad, participant) |>
  as.list()

dat$N_dyads        <- max(dat$dyad)
dat$N_participants <- max(dat$participant)
```

#### Prior predictive
```{r}
mb3_prior <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),

    mu <- a +
      v_dyad[dyad, 1] +
      v_dyad[dyad, 2] * cong +
      b_cond * cond +
      b_interaction * cong * cond,

    a ~ dnorm(6, 0.3),
    b_cond ~ dnorm(0, 0.2),
    b_interaction ~ dnorm(0, 0.2),

    transpars > matrix[N_dyads, 2]:v_dyad <-
      compose_noncentered(sigma_dyad, L_Rho_dyad, z_dyad),
    matrix[2, N_dyads]:z_dyad ~ normal(0, 1),
    cholesky_factor_corr[2]:L_Rho_dyad ~ lkj_corr_cholesky(2),
    vector[2]:sigma_dyad ~ dexp(5),

    sigma ~ dexp(1)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  sample_prior = TRUE
)

prior <- extract.prior(mb3_prior, n = 10000)
saveRDS(mb3_prior, file = "model_and_samples/mb3_prior_model.rds")
saveRDS(prior, file = "model_and_samples/mb3_prior.rds")
```


```{r}
mb3_prior <- readRDS("model_and_samples/mb3_prior_model.rds")
prior <- readRDS("model_and_samples/mb3_prior.rds")
```

We do a prior predictive check:
```{r}
# prior$v_dyad has dimensions [samples, dyads, 2]
# column 1 = intercepts, column 2 = cong slopes
# use mean across dyads for the predictive check

b_cong_prior <- apply(prior$v_dyad[,,2], 1, mean)

rt_ic_as <- exp(prior$a)
rt_c_as  <- exp(prior$a + b_cong_prior)
rt_ic_s  <- exp(prior$a + prior$b_cond)
rt_c_s   <- exp(prior$a + prior$b_cond + b_cong_prior + prior$b_interaction)

par(mfrow = c(2, 2))
dens(rt_ic_as, main = "Incongruent / Async", xlab = "RT (ms)", xlim = c(0, 2000))
dens(rt_c_as,  main = "Congruent / Async",   xlab = "RT (ms)", xlim = c(0, 2000))
dens(rt_ic_s,  main = "Incongruent / Sync",  xlab = "RT (ms)", xlim = c(0, 2000))
dens(rt_c_s,   main = "Congruent / Sync",    xlab = "RT (ms)", xlim = c(0, 2000))
par(mfrow = c(1, 1))

delta_prior <- (rt_c_s - rt_ic_s) - (rt_c_as - rt_ic_as)

data.frame(value = delta_prior) |>
  ggplot(aes(x = value)) +
  geom_density(fill = "forestgreen", colour = "forestgreen", alpha = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  labs(x = "SSE difference: Sync - Async (ms)", y = "Density",
       title = "Prior: SSE Difference (ms scale)")
```

Notice that the prior straddles 0. This is fine, as we find both positive and negative effects plausible a priori.

#### Model fit
```{r}
mb3 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),

    mu <- a +
      v_dyad[dyad, 1] +
      v_dyad[dyad, 2] * cong +
      b_cond * cond +
      b_interaction * cong * cond,

    a ~ dnorm(6, 0.3),
    b_cond ~ dnorm(0, 0.2),
    b_interaction ~ dnorm(0, 0.2),

    transpars > matrix[N_dyads, 2]:v_dyad <-
      compose_noncentered(sigma_dyad, L_Rho_dyad, z_dyad),
    matrix[2, N_dyads]:z_dyad ~ normal(0, 1),
    cholesky_factor_corr[2]:L_Rho_dyad ~ lkj_corr_cholesky(2),
    vector[2]:sigma_dyad ~ dexp(5),

    sigma ~ dexp(1)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  iter = 2000,
  log_lik = TRUE
)

post <- extract.samples(mb3, n = 10000)
saveRDS(mb3, file = "model_and_samples/mb3.rds")
saveRDS(post, file = "model_and_samples/mb3_post.rds")
```

```{r}
mb3 <- readRDS("model_and_samples/mb3.rds")
post <- read_rds("model_and_samples/mb3_post.rds")
```

```{r}
precis(mb3, 2)
```

```{r}
# Compute delta SSE on ms scale
b_cong_post <- apply(post$v_dyad[,,2], 1, mean)  # mean cong slope across dyads

rt_ic_as <- exp(post$a)
rt_c_as  <- exp(post$a + b_cong_post)
rt_ic_s  <- exp(post$a + post$b_cond)
rt_c_s   <- exp(post$a + post$b_cond + b_cong_post + post$b_interaction)

sse_async <- rt_c_as - rt_ic_as
sse_sync  <- rt_c_s  - rt_ic_s
delta_post <- sse_sync - sse_async

# Prior vs posterior: b_interaction (log scale)
data.frame(
  value        = c(prior$b_interaction, post$b_interaction),
  Distribution = c(rep("Prior",     length(prior$b_interaction)),
                   rep("Posterior", length(post$b_interaction)))
) |>
  ggplot(aes(x = value, fill = Distribution, colour = Distribution)) +
  geom_density(alpha = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_fill_manual(values   = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  scale_colour_manual(values = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  labs(x = "b_interaction", y = "Density",
       title = "Prior vs Posterior: Interaction Effect (log scale)")

# Prior vs posterior: SSE difference (ms scale)
b_cong_prior <- apply(prior$v_dyad[,,2], 1, mean)

rt_ic_as_pr <- exp(prior$a)
rt_c_as_pr  <- exp(prior$a + b_cong_prior)
rt_ic_s_pr  <- exp(prior$a + prior$b_cond)
rt_c_s_pr   <- exp(prior$a + prior$b_cond + b_cong_prior + prior$b_interaction)

delta_prior <- (rt_c_s_pr - rt_ic_s_pr) - (rt_c_as_pr - rt_ic_as_pr)

data.frame(
  value        = c(delta_prior, delta_post),
  Distribution = c(rep("Prior",     length(delta_prior)),
                   rep("Posterior", length(delta_post)))
) |>
  ggplot(aes(x = value, fill = Distribution, colour = Distribution)) +
  geom_density(alpha = 0.4) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_fill_manual(values   = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  scale_colour_manual(values = c("Prior" = "forestgreen", "Posterior" = "steelblue")) +
  labs(x = "SSE difference: Sync - Async (ms)", y = "Density",
       title = "Prior vs Posterior: SSE Difference (ms scale)")

# Summary statistics
cat("Posterior mean SSE difference:", round(mean(delta_post), 2), "ms\n")
cat("89% PI:", round(PI(delta_post, prob = 0.89), 2), "\n")
cat("P(delta > 0):", round(mean(delta_post > 0), 3), "\n")
```

```{r}
compare(mb1, mb2, mb3)
```

```{r}
PSIS(mb3, pointwise = TRUE) |>
  as.data.frame() |>
  filter(k > 0.5) |>
  nrow()
PSIS(mb1, pointwise = TRUE) |> as.data.frame() |> filter(k > 0.5) |> nrow()
PSIS(mb2, pointwise = TRUE) |> as.data.frame() |> filter(k > 0.5) |> nrow()
```
```{r}
mb2_psis <- PSIS(mb2, pointwise = TRUE) |> as.data.frame()
problematic <- which(mb2_psis$k > 0.5)
problematic        # indices
mb2_psis[problematic, ]  # their k values and lppd contributions
```

### Posterior predictives
```{r}
# Observed RT
obs_rt <- dat$rt

# mb1: Normal likelihood
sim_mb1 <- sim(mb1, n = 1000)
mb1_mean <- apply(sim_mb1, 2, mean)
mb1_pi   <- apply(sim_mb1, 2, PI, prob = 0.89)

# mb2: Lognormal, fixed slopes
sim_mb2 <- sim(mb2, n = 1000)
mb2_mean <- apply(sim_mb2, 2, mean)
mb2_pi   <- apply(sim_mb2, 2, PI, prob = 0.89)

# mb3: Lognormal, varying slopes
sim_mb3 <- sim(mb3, n = 1000)
mb3_mean <- apply(sim_mb3, 2, mean)
mb3_pi   <- apply(sim_mb3, 2, PI, prob = 0.89)
```

```{r}
# Combine into long format for ggplot
n_obs <- length(obs_rt)
n_sim <- 50  # number of simulated draws to plot

# Sample rows from sim matrices for plotting
set.seed(42)
rows <- sample(1:nrow(sim_mb1), n_sim)

# Build data frames
make_sim_df <- function(sim_mat, rows, model_name) {
  do.call(rbind, lapply(rows, function(i) {
    data.frame(rt = sim_mat[i, ], model = model_name, draw = i)
  }))
}

sim_df <- rbind(
  make_sim_df(sim_mb1, rows, "mb1: Normal"),
  make_sim_df(sim_mb2, rows, "mb2: Lognormal"),
  make_sim_df(sim_mb3, rows, "mb3: Lognormal + varying slopes")
)

obs_df <- data.frame(rt = obs_rt)

# Plot
ggplot() +
  geom_density(
    data = sim_df,
    aes(x = rt, group = interaction(draw, model), colour = model),
    alpha = 0.05, linewidth = 0.3
  ) +
  geom_density(
    data = sim_df |> mutate(obs_rt = rep(obs_rt, n_sim * 3)),
    aes(x = obs_rt),
    colour = "black", linewidth = 1
  ) +
  facet_wrap(~ model, ncol = 1, scales = "free_y") +
  scale_colour_manual(values = c(
    "mb1: Normal"                      = "steelblue",
    "mb2: Lognormal"                   = "forestgreen",
    "mb3: Lognormal + varying slopes"  = "tomato"
  )) +
  coord_cartesian(xlim = c(0, 1000)) +
  labs(x = "RT (ms)", y = "Density",
       title = "Posterior Predictive Check",
       subtitle = "Coloured lines = simulated draws, Black line = observed") +
  guides(colour = "none")
```

```{r}
# One panel per model, observed added separately per facet
obs_df_rep <- data.frame(
  rt    = rep(obs_rt, 3),
  model = rep(c("mb1: Normal", 
                "mb2: Lognormal", 
                "mb3: Lognormal + varying slopes"), 
              each = n_obs)
)

ggplot() +
  geom_density(
    data = sim_df,
    aes(x = rt, group = interaction(draw, model)),
    colour = "steelblue", alpha = 0.05, linewidth = 0.3
  ) +
  geom_density(
    data = obs_df_rep,
    aes(x = rt),
    colour = "black", linewidth = 1
  ) +
  facet_wrap(~ model, ncol = 1, scales = "free_y") +
  coord_cartesian(xlim = c(0, 1000)) +
  labs(x = "RT (ms)", y = "Density",
       title = "Posterior Predictive Check",
       subtitle = "Blue lines = simulated draws, Black line = observed") +
  theme_minimal()
```

