---
title: "Analysis Script for Social and Cultural Dynamics 2026"
author: "Andrea Wiwe Kjær-Hansen, Mikkel Krøjer Svendsen"
format: html
---

# R setup

```{r, message=FALSE}
# Packages
library(rethinking)
library(dagitty)
library(tidyverse)
library(lmerTest)

# Theme settings
theme_set(theme_minimal())

# data directory
source("data_dir.R")
```

# DAG

```{r}
dag <- dagitty(
  'dag {
  synchrony_condition [exposure]
  rt [outcome]
  co_representation [latent]

  synchrony_condition -> co_representation
  synchrony_condition -> rt

  co_representation -> rt

  congruency -> rt
  task_type -> rt
  task_type -> co_representation

  trial_order -> rt
  participant -> rt
}'
)

coordinates(dag) <- list(
  x = c(
    synchrony_condition = 1,
    task_type = 1,
    congruency = 1,
    co_representation = 3,
    rt = 5,
    trial_order = 4,
    participant = 4
  ),
  y = c(
    synchrony_condition = 3,
    task_type = 5,
    congruency = 1,
    co_representation = 3,
    rt = 3,
    trial_order = 1,
    participant = 5
  )
)

plot(dag)
```
```{r}
# What do we need to adjust for to identify the effect of synchrony on RT?
adjustmentSets(dag, exposure = "synchrony_condition", outcome = "rt")

# Are there any testable implications (conditional independencies)?
impliedConditionalIndependencies(dag)
```


# Reading files

```{r}
# read files helper
read_files <- function(pattern) {
  list.files(
    data_dir,
    pattern = pattern,
    full.names = TRUE
  ) |>
    map_dfr(~ read.delim2(.x, sep = ","))
}
```

```{r}
# read files
d_simon <- read_files("simon") |>
  mutate(
    rt_ms = as.numeric(ifelse(rt_ms == "", NA, rt_ms)),
    response_key = ifelse(response_key == "", NA, response_key),
    side_congruent = (shape == "square" & side == "left") |
      (shape == "diamond" & side == "right")
  )

d_tapping <- read_files("tapping")

d_info <- read_files("group_info") |> 
  mutate(
    gender = ifelse(dyad_id == 4 & participant == 2, "Male", gender)
  )
```

# Descriptive statistics and data cleaning
## Participants
```{r}
age <- d_info |>
  summarize(
    age_mean = mean(age),
    age_sd = sd(age)
  )

gender <- d_info |>
  count(gender) |>
  mutate(
    prop = n / sum(n)
  )

condition <- d_tapping |>
  distinct(dyad_id, .keep_all = T) |>
  count(condition, name = "n_dyads")

print(age)
print(gender)
print(condition)
```

## Calculating the Social Simon Effect (SSE)

What is the SSE? In the joint go/no-go Simon task, each participant only responds to one shape. 

The Simon Effect shouldn't exist individually. BUT it often does appear as if participants mentally represent their partner's role. 

So, the SSE: Faster/more accurate responses when the stimulus appears on your response-congruent side, even though congruence is only meaningful relative to the dyad.

$$\text{SSE} = \text{RT}_{\text{congruent}} − \text{RT}_{\text{incongruent}}$$

First, add condition column to data

```{r}
# add condition to d_simon
d_simon <- d_simon |>
  left_join(d_tapping |> distinct(dyad_id, condition), by = "dyad_id")

str(d_simon)
```

We only look at trials that are correct go trials (`responded == 1` and `correct == 1`)

```{r}
d <- d_simon |>
  filter(responded == 1 & correct == 1) |>
  mutate(participant_id = paste(dyad_id, participant, sep = "_")) |>
  filter(
    (participant == 1 & shape == "square") |
      (participant == 2 & shape == "diamond")
  )

str(d)
```

RT summary by condition x task type x congruency
```{r}
rt_summary <- d |>
  group_by(condition, task, side_congruent) |>
  summarise(
    mean_rt = mean(rt_ms, na.rm = TRUE),
    sd_rt = sd(rt_ms, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

rt_summary
```

SSE: congruency effect (incongruent - congruent) by condition x task type
```{r}
congruency_effect <- d |>
  group_by(condition, task, side_congruent) |>
  summarise(mean_rt = mean(rt_ms, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = side_congruent, values_from = mean_rt) |>
  rename(congruent = `TRUE`, incongruent = `FALSE`) |>
  mutate(congruency_effect = incongruent - congruent)

congruency_effect
```
The asynchrony condition looks as expected. There's a small congruency effect in the individual task (0.6 ms) and a larger one in the joint task (18 ms). BUT in the synchrony condition, individuals are slower for congruent trials than incongruent trials. Goes against the Simon effect. This may just be a noisy grand average... worth investigating.

SSE: difference in congruency effect between joint and individual.
```{r}
sse_summary <- congruency_effect |>
  select(condition, task, congruency_effect) |>
  pivot_wider(names_from = task, values_from = congruency_effect) |>
  mutate(SSE = joint - individual)

sse_summary
```

SSE *is* slightly larger for sync than async, though not by a lot.  

This table gives one SSE value per synchrony condition. Maybe good for reporting?

## Visualizations - OLD, PLEASE REVISE!
RT distributions (overlaid)
```{r}
d_joint |>
  mutate(
    side_congruent = ifelse(side_congruent, "Congruent", "Incongruent"),
    condition = str_to_title(condition)
  ) |>
  ggplot(aes(x = rt_ms, fill = side_congruent)) +
  geom_density(alpha = 0.5) +
  facet_wrap(vars(condition)) +
  labs(
    x = "Reaction Time (ms)",
    y = "Density",
    fill = "Congruence",
    title = "RT Distributions by Condition and Congruence"
  )
```
RT distributions (faceted)
```{r}
d_joint |>
  mutate(
    side_congruent = ifelse(side_congruent, "Congruent", "Incongruent"),
    condition = str_to_title(condition)
  ) |>
  ggplot(aes(x = rt_ms, fill = side_congruent)) +
  geom_density(alpha = 0.5) +
  facet_wrap(vars(side_congruent, condition)) +
  labs(
    x = "Reaction Time (ms)",
    y = "Density",
    fill = "Congruence",
    title = "RT Distributions by Condition and Congruence"
  )
```

SSE distributions (condition mean and individual dyads)
```{r}
dyad_sse |>
  mutate(condition = str_to_title(condition)) |>
  ggplot(aes(x = condition, y = SSE, fill = condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_boxplot(alpha = 0.5) +
  geom_jitter(width = 0.05, size = 3, alpha = .5) +
  labs(
    x = "Condition",
    y = "Social Simon Effect (ms)\n[Incongruent − Congruent RT]",
    title = "SSE per Dyad by Condition"
  ) +
  theme(legend.position = "none")
```

Congruence effect per dyad by condition
```{r}
d_joint |>
  group_by(dyad_id, condition, side_congruent) |>
  summarise(mean_rt = mean(rt_ms), .groups = "drop") |>
  mutate(
    side_congruent = ifelse(side_congruent, "Congruent", "Incongruent"),
    condition = str_to_title(condition)
  ) |>
  ggplot(aes(
    x = side_congruent,
    y = mean_rt,
    group = dyad_id,
    colour = factor(dyad_id)
  )) +
  geom_line() +
  geom_point(size = 3) +
  facet_wrap(~condition) +
  labs(
    x = "",
    y = "Mean RT (ms)",
    colour = "Dyad",
    title = "Congruence Effect per Dyad by Condition"
  )
```

# Analysis
## Model 0 - intercept only
### Data
```{r}
d_model <- d |>
  mutate(
    congruent = as.integer(side_congruent), # 0 = incongruent, 1 = congruent
    joint = as.integer(task == "joint"), # 0 = individual, 1 = joint
    sync = as.integer(condition == "synchrony"), # 0 = async, 1 = sync
    dyad_id = as.integer(as.factor(dyad_id)),
    participant_id = as.integer(as.factor(participant_id))
  )

dat <- list(
  rt = d_model$rt_ms,
  congruent = d_model$congruent,
  joint = d_model$joint,
  sync = d_model$sync,
  dyad = d_model$dyad_id,
  pid = d_model$participant_id
)
```

### Prior predictive
For a lognormal: if log(RT) ~ Normal(mu, sigma) then RT is in ms. Typical RT ~400ms, so log(400) ≈ 6.0. We want priors that produce plausible RT distributions.

```{r}
n_samples <- 2000
prior_mu <- rnorm(n_samples, mean = 6, sd = .5)
prior_sigma <- rexp(n_samples, rate = 1)
prior_rt <- rlnorm(n_samples, meanlog = prior_mu, sdlog = prior_sigma)

# Plot
tibble(rt = prior_rt) |>
  filter(rt < 5000) |> # trim extreme tails for visualisation
  ggplot(aes(x = rt)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check — Model 0",
    subtitle = "Red dashed lines: plausible RT range (150–1800 ms)",
    x = "RT (ms)",
    y = "Density"
  )
```


### Model fit
```{r}
m0 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a,
    a ~ dnorm(6, .5),
    sigma ~ dexp(1)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
trankplot(m0)
```


```{r}
precis(m0)
```
### Posterior predictive check
sim() draws from the posterior predictive distribution (includes sigma)
link() draws mu only. Use sim() here since we want full RT distributions:
```{r}
rt_sim0 <- sim(m0, n = n_samples)

# One simulated RT per observation per posterior draw
# colMeans gives the average posterior predictive per observation,
# but for a density plot we want the full distribution —
# so flatten the entire matrix instead

rt_ppc0 <- as.vector(rt_sim0)

tibble(
  rt     = c(dat$rt, rt_ppc0),
  source = c(rep("Observed", length(dat$rt)),
             rep("Posterior predictive", length(rt_ppc0)))
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Posterior predictive check — Model 0",
    x = "RT (ms)", y = "Density", fill = ""
  )
```
### Prior-posterior comparison
```{r}
n_post <- length(post0$a)

prior_post0 <- tibble(
  prior__a        = rnorm(n_post, 6.0, 0.5),
  posterior__a    = post0$a,
  prior__sigma    = rexp(n_post, 1),
  posterior__sigma = post0$sigma
) |>
  pivot_longer(everything(),
               names_to  = c("source", "parameter"),
               names_sep = "__")

# Compute posterior range per parameter for zooming
zoom <- prior_post0 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001),
            xmax = quantile(value, 0.999))

prior_post0 |>
  left_join(zoom, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison — Model 0",
    x = "", y = "Density", fill = ""
  )
```

## Model 1 - main effects
### Data
```{r}
# still dat
```

### Prior predictive
```{r}
n_post <- 2000  # align with 4 chains x 500 draws

prior_a <- rnorm(n_post, 6.0, 0.5)
prior_b_cong <- rnorm(n_post, 0, 0.5)
prior_b_joint <- rnorm(n_post, 0, 0.5)
prior_b_sync <- rnorm(n_post, 0, 0.5)
prior_sigma <- rexp(n_post, 1)

# Compute prior mu for each combination of predictors (reference cell: 0/0/0)
prior_mu <- prior_a +
  prior_b_cong * 1 + # congruent
  prior_b_joint * 1 + # joint
  prior_b_sync * 1 # sync

prior_rt <- rlnorm(n_post, meanlog = prior_mu, sdlog = prior_sigma)

tibble(rt = prior_rt) |>
  filter(rt < 5000) |>
  ggplot(aes(x = rt)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check — Model 1",
    subtitle = "Red dashed lines: plausible RT range (150–1800 ms)",
    x = "RT (ms)",
    y = "Density"
  )
```

### Model fit
```{r}
m1 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong * congruent +
      b_joint * joint +
      b_sync * sync,
    a ~ dnorm(6.0, 0.5),
    b_cong ~ dnorm(0, 0.5),
    b_joint ~ dnorm(0, 0.5),
    b_sync ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = dat,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
precis(m1)
```

### Posterior predictive check
```{r}
rt_sim1 <- sim(m1)
rt_ppc1 <- as.vector(rt_sim1)

tibble(
  rt = c(dat$rt, rt_ppc1),
  source = c(
    rep("Observed", length(dat$rt)),
    rep("Posterior predictive", length(rt_ppc1))
  )
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Posterior predictive check — Model 1",
    x = "RT (ms)",
    y = "Density",
    fill = ""
  )
```

### Prior-posterior comparison
```{r}
n_post <- length(post1$a)

prior_post1 <- tibble(
  prior__a          = rnorm(n_post, 6.0, 0.5),
  posterior__a      = post1$a,
  prior__cong       = rnorm(n_post, 0, 0.5),
  posterior__cong   = post1$b_cong,
  prior__joint      = rnorm(n_post, 0, 0.5),
  posterior__joint  = post1$b_joint,
  prior__sync       = rnorm(n_post, 0, 0.5),
  posterior__sync   = post1$b_sync,
  prior__sigma      = rexp(n_post, 1),
  posterior__sigma  = post1$sigma
) |>
  pivot_longer(everything(),
               names_to  = c("source", "parameter"),
               names_sep = "__")

zoom1 <- prior_post1 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001), xmax = quantile(value, 0.999))

prior_post1 |>
  left_join(zoom1, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison — Model 1",
    x = "",
    y = "Density",
    fill = ""
  )
```

## Model 2 - main effects + varying intercepts
### Data
### Prior predictive
```{r}
n_post <- 2000

prior_a         <- rnorm(n_post, 6.0, 0.5)
prior_b_cong    <- rnorm(n_post, 0, 0.5)
prior_b_joint   <- rnorm(n_post, 0, 0.5)
prior_b_sync    <- rnorm(n_post, 0, 0.5)
prior_sigma     <- rexp(n_post, 1)
prior_sigma_pid <- rexp(n_post, 1)
prior_sigma_dyad <- rexp(n_post, 1)

# Simulate varying intercepts
prior_u_pid  <- rnorm(n_post, 0, prior_sigma_pid)
prior_u_dyad <- rnorm(n_post, 0, prior_sigma_dyad)

prior_mu <- prior_a +
  prior_b_cong  * 1 +
  prior_b_joint * 1 +
  prior_b_sync  * 1 +
  prior_u_pid +
  prior_u_dyad

prior_rt <- rlnorm(n_post, meanlog = prior_mu, sdlog = prior_sigma)

tibble(rt = prior_rt) |>
  filter(rt < 5000) |>
  ggplot(aes(x = rt)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check — Model 2",
    subtitle = "Red dashed lines: plausible RT range (150–1800 ms)",
    x = "RT (ms)", y = "Density"
  )
```

### Model fit
```{r}
m2 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong  * congruent +
      b_joint * joint +
      b_sync  * sync +
      u_pid[pid] +
      u_dyad[dyad],
    
    # Fixed effects
    a       ~ dnorm(6.0, 0.5),
    b_cong  ~ dnorm(0, 0.5),
    b_joint ~ dnorm(0, 0.5),
    b_sync  ~ dnorm(0, 0.5),
    
    # Varying intercepts
    u_pid[pid]   ~ dnorm(0, sigma_pid),
    u_dyad[dyad] ~ dnorm(0, sigma_dyad),
    
    # Scale parameters
    sigma      ~ dexp(1),
    sigma_pid  ~ dexp(1),
    sigma_dyad ~ dexp(1)
  ),
  data = dat, chains = 4, cores = 4, log_lik = TRUE
)
```

```{r}
precis(m2)
```

### Posterior predictive check
```{r}
rt_sim2 <- sim(m2)
rt_ppc2 <- as.vector(rt_sim2)

tibble(
  rt     = c(dat$rt, rt_ppc2),
  source = c(rep("Observed", length(dat$rt)),
             rep("Posterior predictive", length(rt_ppc2)))
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Posterior predictive check — Model 2",
    x = "RT (ms)", y = "Density", fill = ""
  )
```


### Prior-posterior comparison
```{r}
n_post <- length(post2$a)
prior_post2 <- tibble(
  prior__a           = rnorm(n_post, 6.0, 0.5),
  posterior__a       = post2$a,
  prior__cong        = rnorm(n_post, 0, 0.5),
  posterior__cong    = post2$b_cong,
  prior__joint       = rnorm(n_post, 0, 0.5),
  posterior__joint   = post2$b_joint,
  prior__sync        = rnorm(n_post, 0, 0.5),
  posterior__sync    = post2$b_sync,
  prior__sigma       = rexp(n_post, 1),
  posterior__sigma   = post2$sigma,
  prior__sigma_pid   = rexp(n_post, 1),
  posterior__sigma_pid  = post2$sigma_pid,
  prior__sigma_dyad  = rexp(n_post, 1),
  posterior__sigma_dyad = post2$sigma_dyad
) |>
  pivot_longer(everything(),
               names_to  = c("source", "parameter"),
               names_sep = "__")

zoom2 <- prior_post2 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001),
            xmax = quantile(value, 0.999))

prior_post2 |>
  left_join(zoom2, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison — Model 2",
    x = "", y = "Density", fill = ""
  )
```

## Model 3 - two-way interactions + varying intercepts
### Data
### Prior predictive
```{r}
n_post <- 2000

prior_a              <- rnorm(n_post, 6.0, 0.5)
prior_b_cong         <- rnorm(n_post, 0, 0.5)
prior_b_joint        <- rnorm(n_post, 0, 0.5)
prior_b_sync         <- rnorm(n_post, 0, 0.5)
prior_b_cong_joint   <- rnorm(n_post, 0, 0.5)
prior_b_cong_sync    <- rnorm(n_post, 0, 0.5)
prior_b_joint_sync   <- rnorm(n_post, 0, 0.5)
prior_sigma          <- rexp(n_post, 1)
prior_sigma_pid      <- rexp(n_post, 1)
prior_sigma_dyad     <- rexp(n_post, 1)

prior_u_pid  <- rnorm(n_post, 0, prior_sigma_pid)
prior_u_dyad <- rnorm(n_post, 0, prior_sigma_dyad)

prior_mu <- prior_a +
  prior_b_cong       * 1 +
  prior_b_joint      * 1 +
  prior_b_sync       * 1 +
  prior_b_cong_joint * 1 * 1 +
  prior_b_cong_sync  * 1 * 1 +
  prior_b_joint_sync * 1 * 1 +
  prior_u_pid +
  prior_u_dyad

prior_rt <- rlnorm(n_post, meanlog = prior_mu, sdlog = prior_sigma)

tibble(rt = prior_rt) |>
  filter(rt < 5000) |>
  ggplot(aes(x = rt)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check — Model 3",
    subtitle = "Red dashed lines: plausible RT range (150–1800 ms)",
    x = "RT (ms)", y = "Density"
  )
```

### Model fit
```{r}
m3 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong         * congruent +
      b_joint        * joint +
      b_sync         * sync +
      b_cong_joint   * congruent * joint +
      b_cong_sync    * congruent * sync +
      b_joint_sync   * joint * sync +
      u_pid[pid] +
      u_dyad[dyad],

    # Fixed effects
    a              ~ dnorm(6.0, 0.5),
    b_cong         ~ dnorm(0, 0.5),
    b_joint        ~ dnorm(0, 0.5),
    b_sync         ~ dnorm(0, 0.5),
    b_cong_joint   ~ dnorm(0, 0.5),
    b_cong_sync    ~ dnorm(0, 0.5),
    b_joint_sync   ~ dnorm(0, 0.5),

    # Varying intercepts
    u_pid[pid]   ~ dnorm(0, sigma_pid),
    u_dyad[dyad] ~ dnorm(0, sigma_dyad),

    # Scale parameters
    sigma      ~ dexp(1),
    sigma_pid  ~ dexp(1),
    sigma_dyad ~ dexp(1)
  ),
  data = dat, chains = 4, cores = 4, log_lik = TRUE
)
```

```{r}
precis(m3, depth = 1)
```

### Posterior predictive check
```{r}
rt_sim3 <- sim(m3)
rt_ppc3 <- as.vector(rt_sim3)

tibble(
  rt     = c(dat$rt, rt_ppc3),
  source = c(rep("Observed", length(dat$rt)),
             rep("Posterior predictive", length(rt_ppc3)))
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Posterior predictive check — Model 3",
    x = "RT (ms)", y = "Density", fill = ""
  )
```

### Prior-posterior comparison
```{r}
post3  <- extract.samples(m3)
n_post <- length(post3$a)

prior_post3 <- tibble(
  prior__a                = rnorm(n_post, 6.0, 0.5),
  posterior__a            = post3$a,
  prior__cong             = rnorm(n_post, 0, 0.5),
  posterior__cong         = post3$b_cong,
  prior__joint            = rnorm(n_post, 0, 0.5),
  posterior__joint        = post3$b_joint,
  prior__sync             = rnorm(n_post, 0, 0.5),
  posterior__sync         = post3$b_sync,
  prior__cong_joint       = rnorm(n_post, 0, 0.5),
  posterior__cong_joint   = post3$b_cong_joint,
  prior__cong_sync        = rnorm(n_post, 0, 0.5),
  posterior__cong_sync    = post3$b_cong_sync,
  prior__joint_sync       = rnorm(n_post, 0, 0.5),
  posterior__joint_sync   = post3$b_joint_sync,
  prior__sigma            = rexp(n_post, 1),
  posterior__sigma        = post3$sigma,
  prior__sigma_pid        = rexp(n_post, 1),
  posterior__sigma_pid    = post3$sigma_pid,
  prior__sigma_dyad       = rexp(n_post, 1),
  posterior__sigma_dyad   = post3$sigma_dyad
) |>
  pivot_longer(everything(),
               names_to  = c("source", "parameter"),
               names_sep = "__")

zoom3 <- prior_post3 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001),
            xmax = quantile(value, 0.999))

prior_post3 |>
  left_join(zoom3, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison — Model 3",
    x = "", y = "Density", fill = ""
  )
```

## Model 4 - the whole shebang!
### Data
### Prior predictive
```{r}
n_post <- 2000

prior__a                  <- rnorm(n_post, 6.0, 0.5)
prior__cong               <- rnorm(n_post, 0, 0.5)
prior__joint              <- rnorm(n_post, 0, 0.5)
prior__sync               <- rnorm(n_post, 0, 0.5)
prior__cong_joint         <- rnorm(n_post, 0, 0.5)
prior__cong_sync          <- rnorm(n_post, 0, 0.5)
prior__joint_sync         <- rnorm(n_post, 0, 0.5)
prior__cong_joint_sync    <- rnorm(n_post, 0, 0.5)
prior__sigma              <- rexp(n_post, 1)
prior__sigma_pid          <- rexp(n_post, 1)
prior__sigma_dyad         <- rexp(n_post, 1)

prior_u_pid  <- rnorm(n_post, 0, prior__sigma_pid)
prior_u_dyad <- rnorm(n_post, 0, prior__sigma_dyad)

prior_mu <- prior__a +
  prior__cong            * 1 +
  prior__joint           * 1 +
  prior__sync            * 1 +
  prior__cong_joint      * 1 * 1 +
  prior__cong_sync       * 1 * 1 +
  prior__joint_sync      * 1 * 1 +
  prior__cong_joint_sync * 1 * 1 * 1 +
  prior_u_pid +
  prior_u_dyad

prior_rt <- rlnorm(n_post, meanlog = prior_mu, sdlog = prior__sigma)

tibble(rt = prior_rt) |>
  filter(rt < 5000) |>
  ggplot(aes(x = rt)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = c(150, 1800), linetype = "dashed", colour = "red") +
  labs(
    title = "Prior predictive check — Model 4",
    subtitle = "Red dashed lines: plausible RT range (150–1800 ms)",
    x = "RT (ms)", y = "Density"
  )
```

### Model fit
```{r}
m4 <- ulam(
  alist(
    rt ~ dlnorm(mu, sigma),
    mu <- a +
      b_cong            * congruent +
      b_joint           * joint +
      b_sync            * sync +
      b_cong_joint      * congruent * joint +
      b_cong_sync       * congruent * sync +
      b_joint_sync      * joint * sync +
      b_cong_joint_sync * congruent * joint * sync +
      u_pid[pid] +
      u_dyad[dyad],

    # Fixed effects
    a                 ~ dnorm(6.0, 0.5),
    b_cong            ~ dnorm(0, 0.5),
    b_joint           ~ dnorm(0, 0.5),
    b_sync            ~ dnorm(0, 0.5),
    b_cong_joint      ~ dnorm(0, 0.5),
    b_cong_sync       ~ dnorm(0, 0.5),
    b_joint_sync      ~ dnorm(0, 0.5),
    b_cong_joint_sync ~ dnorm(0, 0.5),

    # Varying intercepts
    u_pid[pid]   ~ dnorm(0, sigma_pid),
    u_dyad[dyad] ~ dnorm(0, sigma_dyad),

    # Scale parameters
    sigma      ~ dexp(1),
    sigma_pid  ~ dexp(1),
    sigma_dyad ~ dexp(1)
  ),
  data = dat, chains = 4, cores = 4, log_lik = TRUE
)
```

```{r}
precis(m4, depth = 1)
```

### Posterior predictive check
```{r}
rt_sim4 <- sim(m4)
rt_ppc4 <- as.vector(rt_sim4)

tibble(
  rt     = c(dat$rt, rt_ppc4),
  source = c(rep("Observed", length(dat$rt)),
             rep("Posterior predictive", length(rt_ppc4)))
) |>
  filter(rt < 2000) |>
  ggplot(aes(x = rt, fill = source)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Posterior predictive check — Model 4",
    x = "RT (ms)", y = "Density", fill = ""
  )
```

### Prior-posterior comparison
```{r}
post4  <- extract.samples(m4)
n_post <- length(post4$a)

prior_post4 <- tibble(
  prior__a                  = rnorm(n_post, 6.0, 0.5),
  posterior__a              = post4$a,
  prior__cong               = rnorm(n_post, 0, 0.5),
  posterior__cong           = post4$b_cong,
  prior__joint              = rnorm(n_post, 0, 0.5),
  posterior__joint          = post4$b_joint,
  prior__sync               = rnorm(n_post, 0, 0.5),
  posterior__sync           = post4$b_sync,
  prior__cong_joint         = rnorm(n_post, 0, 0.5),
  posterior__cong_joint     = post4$b_cong_joint,
  prior__cong_sync          = rnorm(n_post, 0, 0.5),
  posterior__cong_sync      = post4$b_cong_sync,
  prior__joint_sync         = rnorm(n_post, 0, 0.5),
  posterior__joint_sync     = post4$b_joint_sync,
  prior__cong_joint_sync    = rnorm(n_post, 0, 0.5),
  posterior__cong_joint_sync = post4$b_cong_joint_sync,
  prior__sigma              = rexp(n_post, 1),
  posterior__sigma          = post4$sigma,
  prior__sigma_pid          = rexp(n_post, 1),
  posterior__sigma_pid      = post4$sigma_pid,
  prior__sigma_dyad         = rexp(n_post, 1),
  posterior__sigma_dyad     = post4$sigma_dyad
) |>
  pivot_longer(everything(),
               names_to  = c("source", "parameter"),
               names_sep = "__")

zoom4 <- prior_post4 |>
  filter(source == "posterior") |>
  group_by(parameter) |>
  summarise(xmin = quantile(value, 0.001),
            xmax = quantile(value, 0.999))

prior_post4 |>
  left_join(zoom4, by = "parameter") |>
  filter(value >= xmin & value <= xmax) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ parameter, scales = "free") +
  labs(
    title = "Prior-posterior comparison — Model 4",
    x = "", y = "Density", fill = ""
  )
```
## Model comparison
```{r}
comparison <- compare(m0, m1, m2, m3, m4)
comparison
plot(comparison)
```

M3 and M4 are essentially equivalent — dWAIC of 2.0 with a dSE of 0.71 means the difference is real but trivially small. The model weights reflect this: 0.73 vs 0.27. Adding the three-way interaction (M4) does not meaningfully improve predictive accuracy over the two-way interaction model (M3).
M2 is clearly worse than M3/M4 — dWAIC of 14.6 with dSE of 8.98 suggests the two-way interactions are earning their place. The data support including them.
M1 and M0 are far behind — the large dWAIC values confirm that both the fixed effects and the varying intercepts are doing real work.
The pWAIC column is worth noting: M3 has 23.5 effective parameters and M4 has 24.4 — only about 1 effective parameter difference despite adding a full interaction term. This suggests the three-way interaction is not being constrained much by the data, consistent with the wide posterior we saw.
Practical conclusion: M3 is your best model by WAIC. M4 is your theoretically motivated model but adds no predictive value. This is a meaningful result in itself — you can report that the three-way interaction does not improve model fit, and connect that to the wide posterior and small sample as converging evidence that the data are insufficient to estimate the effect of synchrony on the SSE.

## Posterior estimates of m4
```{r}
# Extract samples
post4 <- extract.samples(m4)

# Key parameters: mean, SD, and 89% credible interval

params <- c(
  "b_cong",
  "b_joint",
  "b_sync",
  "b_cong_joint",
  "b_cong_sync",
  "b_joint_sync",
  "b_cong_joint_sync"
)

posterior_summary <- map_dfr(params, function(p) {
  samples <- post4[[p]]
  tibble(
    parameter = p,
    mean = mean(samples),
    sd = sd(samples),
    lower_89 = quantile(samples, 0.055),
    upper_89 = quantile(samples, 0.945)
  )
})

posterior_summary
```

Remember this is in log-RT. 

b_cong_joint (-0.057, 89% CI [-0.107, -0.006]): This is the SSE — the congruency effect is more negative in the joint task than the individual task. CI excludes zero, so there is credible evidence for the SSE in the asynchrony condition. The negative sign means incongruent trials are slower relative to congruent in the joint task.
b_cong_sync (0.030, 89% CI [-0.016, 0.077]): Weak suggestion that the congruency effect differs slightly by condition in the individual task. CI includes zero.

b_cong_joint_sync (-0.014, 89% CI [-0.082, 0.050]): The SSE does not credibly differ between synchrony and asynchrony conditions. The mean is slightly negative — suggesting the SSE might be marginally smaller in the synchrony condition — but the CI is wide and comfortably includes zero. This is your main null result.

## Contrasts - SSE by condition
### On log scale
```{r}
# SSE in each condition (on log-RT scale)
SSE_async <- post4$b_cong_joint
SSE_sync  <- post4$b_cong_joint + post4$b_cong_joint_sync

# Contrast: difference in SSE between conditions
SSE_contrast <- SSE_sync - SSE_async

# Summarise
contrast_summary <- tibble(
  parameter     = c("SSE_async", "SSE_sync", "SSE_contrast"),
  mean          = c(mean(SSE_async), mean(SSE_sync), mean(SSE_contrast)),
  sd            = c(sd(SSE_async), sd(SSE_sync), sd(SSE_contrast)),
  lower_89      = c(quantile(SSE_async, 0.055),
                    quantile(SSE_sync,  0.055),
                    quantile(SSE_contrast, 0.055)),
  upper_89      = c(quantile(SSE_async, 0.945),
                    quantile(SSE_sync,  0.945),
                    quantile(SSE_contrast, 0.945))
)

contrast_summary

# Plot contrast distribution
tibble(contrast = SSE_contrast) |>
  ggplot(aes(x = contrast)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  labs(
    title = "Posterior contrast: SSE synchrony vs. asynchrony",
    subtitle = "Negative = smaller SSE in synchrony condition",
    x = "SSE_sync - SSE_async (log-RT scale)", y = "Density"
  )
```

SSE_async (-0.057, 89% CI [-0.107, -0.006]): The SSE is credibly present in the asynchrony condition — the CI excludes zero. Incongruent trials are slower than congruent in the joint task relative to the individual task.
SSE_sync (-0.070, 89% CI [-0.113, -0.028]): The SSE is also credibly present in the synchrony condition, and actually slightly larger in magnitude than in the asynchrony condition. CI excludes zero more cleanly here.
SSE_contrast (-0.014, 89% CI [-0.082, 0.050]): The difference between conditions is negligible and highly uncertain. The CI straddles zero almost symmetrically, and the plot shows a distribution centered essentially on zero with substantial spread in both directions. There is no credible evidence that synchrony modulates the size of the SSE.
The plot tells the same story cleanly — the posterior contrast is centered just left of zero with a long, roughly symmetric spread. Most of the mass is between -0.10 and +0.05, which on the log-RT scale is a very small range.
Summary for your paper: Both conditions show a credible SSE, but the contrast between them is not credibly different from zero. This is a clean null result on the primary hypothesis, with the honest caveat that the sample size gives very little power to detect a difference at the dyad level.

### Transformed to ms scale
```{r}
# Baseline: reference cell (async, individual, incongruent)
baseline <- exp(post4$a)

# SSE in each condition: exp(a + SSE) - exp(a)
SSE_async_ms <- exp(post4$a + SSE_async) - baseline
SSE_sync_ms  <- exp(post4$a + SSE_sync)  - baseline
SSE_contrast_ms <- SSE_sync_ms - SSE_async_ms

ms_summary <- tibble(
  parameter = c("SSE_async_ms", "SSE_sync_ms", "SSE_contrast_ms"),
  mean      = c(mean(SSE_async_ms), mean(SSE_sync_ms), mean(SSE_contrast_ms)),
  sd        = c(sd(SSE_async_ms), sd(SSE_sync_ms), sd(SSE_contrast_ms)),
  lower_89  = c(quantile(SSE_async_ms, 0.055),
                quantile(SSE_sync_ms,  0.055),
                quantile(SSE_contrast_ms, 0.055)),
  upper_89  = c(quantile(SSE_async_ms, 0.945),
                quantile(SSE_sync_ms,  0.945),
                quantile(SSE_contrast_ms, 0.945))
)

ms_summary

# Plot contrast in ms
tibble(contrast_ms = SSE_contrast_ms) |>
  ggplot(aes(x = contrast_ms)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  labs(
    title = "Posterior contrast: SSE synchrony vs. asynchrony (ms)",
    subtitle = "Negative = smaller SSE in synchrony condition",
    x = "SSE_sync - SSE_async (ms)", y = "Density"
  )
```

SSE_async (-19.9 ms, 89% CI [-36.8, -2.2]): A credible SSE of roughly 20 ms in the asynchrony condition — incongruent trials are about 20 ms slower than congruent in the joint task relative to the individual task. CI excludes zero.
SSE_sync (-24.9 ms, 89% CI [-40.1, -10.2]): A credible SSE of roughly 25 ms in the synchrony condition. Actually slightly larger in magnitude than the asynchrony condition, and estimated more precisely — CI excludes zero more cleanly.
SSE_contrast (-4.9 ms, 89% CI [-28.6, 17.3]): The difference between conditions is tiny and completely uncertain. The CI spans roughly 46 ms and straddles zero almost symmetrically. The plot confirms this — a broad, roughly symmetric distribution centered just left of zero.
For your results section, these are the numbers to report. The story is:

The SSE was present in both conditions, with posterior means of approximately 20 ms (asynchrony) and 25 ms (synchrony)
The contrast between conditions was -4.9 ms (89% CI [-28.6, 17.3]), providing no credible evidence that interpersonal synchrony modulated the size of the SSE